# Top 30 `r region(2)` {.tabset .tabset-fade .tabset-pills}

## Total Cases

```{r, message=FALSE}
top30(my.corona, 'confirmed')
```

## Total Deaths

```{r, message=FALSE}
top30(my.corona, 'death')
```

## Data Cases

```{r}
my.corona %>% top30.data('confirmed') %>% datatable
```

## Data Deaths

```{r}
my.corona %>% top30.data('death') %>% datatable
```

# Confirmed cases / Deaths {.tabset .tabset-fade .tabset-pills}

## Confirmed

```{r}
ommit.start(my.corona, 'confirmed', params$confirmed.start, filter.states = countries)
```

## Deaths

```{r}
ommit.start(my.corona, 'death', params$confirmed.start, filter.states = countries)
```

## Confirmed *(log2 scale)*

```{r}
ommit.start(my.corona, 'confirmed', params$confirmed.start, filter.states = countries, log2.flag = TRUE)
```

## Deaths *(log2 scale)*

```{r}
ommit.start(my.corona, 'death', params$confirmed.start, filter.states = countries, log2.flag = TRUE)
```

## Data for Confirmed

```{r}
ommit.data(my.corona, 'confirmed') %>% datatable
```

## Data for Deaths

```{r}
ommit.data(my.corona, 'confirmed') %>% datatable
```

# Confirmed cases / Deaths *(ommiting early start of epidemic)*

## Starting point after exceeding `r params$confirmed.start` cases or `r params$death.start` deaths {.tabset .tabset-fade .tabset-pills}

The visualizations presented below have been time-shifted to the day when each `r region(1, TRUE)` exceeded `r params$confirmed.start` cases or `r params$death.start` deaths *(depending on the visualization)*.

It is an arbitrary number, but allows to have a more accurate view of the progression as the first cases are quite sporadic and make the plots more complex.

**Important note**: All plots are in the logarithm scale *(base 2)* as it allows to better visualize when cases duplicate.

It is used because covid-19 cases grow very fast, and this way it becomes easier to read over time the number of cases. Always look at the y axis.

If you do not understand logarythm scale, I recommend this [video](https://www.youtube.com/watch?v=Kas0tIxDvrg).

### Cases  

```{r}
ommit.start(after.100.dat$all, 'confirmed', params$confirmed.start, filter.states = countries)
```

### Deaths

```{r}
ommit.start(after.100.dat$all, 'death', params$death.start, filter.states = countries)
```

### Cases *(log2 scale)*

```{r}
ommit.start(after.100.dat$all, 'confirmed', params$confirmer.start, filter.states = countries, log2.flag = TRUE)
```

### Deaths *(log2 scale)*

```{r}
ommit.start(after.100.dat$all, 'death', params$death.start, filter.states = countries, log2.flag = TRUE)
```

### Data for Cases

```{r}
after.100.dat$confirmed %>% datatable
```

### Data for Deaths

```{r}
after.100.dat$death %>% datatable()
```

## Starting point after exceeding `r params$confirmed.start` cases or `r params$death.start` deaths *(per 100k population)* {.tabset .tabset-fade .tabset-pills}

Showing population for plotted `r region(2, TRUE)` This becomes relevant for the next plots.

### Cases

```{r}
ommit.start(after.100.dat$all, 'confirmed', params$confirmed.start, filter.states = countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

### Deaths 

```{r}
ommit.start(after.100.dat$all, 'death', params$confirmer.start, filter.states = countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

### Cases *(log2 scale)*

```{r, warning=FALSE}
ommit.start(after.100.dat$all, 'confirmed', params$confirmer.start, filter.states = countries, log2.flag = TRUE, per.100k.flag = TRUE)
```

### Deaths *(log2 scale)*

```{r, warning=FALSE}
ommit.start(after.100.dat$all, 'death', params$confirmer.start, filter.states = countries, log2.flag = TRUE, per.100k.flag = TRUE)
```

### Data for Cases

```{r}
after.100.dat$confirmed %>% datatable
```

### Data for Deaths

```{r}
after.100.dat$death %>% datatable
```

# Last `r params$last.days` days

## Cases/Deaths by day in Last `r params$last.days` days {.tabset .tabset-fade .tabset-pills}

Instead of showing the evolution of the cases/deaths since a `r region(1, TRUE)` exceeded `r params$confirmed.start`/`r params$death.start` cases, it insteads shows the last 12 days.

The first data point is the number of new cases that day.

### Cases 

```{r}
last.days(last.days.dat, 'confirmed', params$last.days, countries, log2.flag = FALSE, per.100k.flag = FALSE, new.flag = TRUE)
```

### Deaths

```{r, warning=FALSE}
last.days(last.days.dat, 'death', params$last.days, countries, log2.flag = FALSE, per.100k.flag = FALSE, new.flag = TRUE)
```

### Cases *(log2 scale)*

```{r, message=FALSE, warning=FALSE}
last.days(last.days.dat, 'confirmed', params$last.days, countries, log2.flag = TRUE, per.100k.flag = FALSE, new.flag = TRUE)
```

### Deaths

```{r, warning=FALSE}
last.days(last.days.dat, 'death', params$last.days, countries, log2.flag = TRUE, per.100k.flag = FALSE, new.flag = TRUE)
```

### Data for Cases

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(last.days.dat, 'confirmed') %>% datatable
```

### Data for Deaths

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(last.days.dat, 'death') %>% datatable
```

## Cases/Deaths by day in Last `r params$last.days` days (per 100k population) {.tabset .tabset-fade .tabset-pills}

Instead of showing the evolution of the cases/deaths since a `r region(1, TRUE)` exceeded `r params$confirmed.start`/`r params$death.start` cases, it insteads shows the last 12 days.

The first data point is the number of new cases that day.

### Cases 

```{r}
last.days(last.days.dat, 'confirmed', params$last.days, countries, log2.flag = FALSE, per.100k.flag = TRUE, new.flag = TRUE)
```

### Deaths

```{r, warning=FALSE}
last.days(last.days.dat, 'death', params$last.days, countries, log2.flag = FALSE, per.100k.flag = TRUE, new.flag = TRUE)
```

### Cases *(log2 scale)*

```{r, message=FALSE, warning=FALSE}
last.days(last.days.dat, 'confirmed', params$last.days, countries, log2.flag = TRUE, per.100k.flag = TRUE, new.flag = TRUE)
```

### Deaths

```{r, warning=FALSE}
last.days(last.days.dat, 'death', params$last.days, countries, log2.flag = TRUE, per.100k.flag = TRUE, new.flag = TRUE)
```

### Data for Cases

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(last.days.dat, 'confirmed') %>% datatable
```

### Data for Deaths

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(last.days.dat, 'death') %>% datatable
```

## Total Cases/Deaths in the last `r params$last.days` days {.tabset .tabset-fade .tabset-pills}

Instead of showing the evolution of the cases/deaths since a `r region(1, TRUE)` exceeded `r params$confirmed.start`/`r params$death.start` cases, it insteads shows the last 12 days.

The first data point is the number of new cases that day.

### Cases 

```{r}
last.days(last.days.dat, 'confirmed', params$last.days, countries, log2.flag = FALSE, per.100k.flag = FALSE)
```

### Deaths

```{r, warning=FALSE}
last.days(last.days.dat, 'death', params$last.days, countries, log2.flag = FALSE, per.100k.flag = FALSE)
```

### Cases *(log2 scale)*

```{r, message=FALSE, warning=FALSE}
last.days(last.days.dat, 'confirmed', params$last.days, countries, log2.flag = TRUE, per.100k.flag = FALSE)
```

### Deaths

```{r, warning=FALSE}
last.days(last.days.dat, 'death', params$last.days, countries, log2.flag = TRUE, per.100k.flag = FALSE)
```

### Data for Cases

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(last.days.dat, 'confirmed', countries) %>% datatable
```

### Data for Deaths

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(last.days.dat, 'death', countries) %>% datatable
```

## Total Cases/Deaths in the last `r params$last.days` days *(per 100k population)* {.tabset .tabset-fade .tabset-pills}

Instead of showing the evolution of the cases/deaths since a `r region(1, TRUE)` exceeded `r params$confirmed.start`/`r params$death.start` cases, it insteads shows the last 12 days.

The first data point is the number of new cases that day.

### Cases 

```{r}
last.days(last.days.dat, 'confirmed', params$last.days, countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

### Deaths

```{r, warning=FALSE}
last.days(last.days.dat, 'death', params$last.days, countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

### Cases *(log2 scale)*

```{r, message=FALSE, warning=FALSE}
last.days(last.days.dat, 'confirmed', params$last.days, countries, log2.flag = TRUE, per.100k.flag = TRUE)
```

### Deaths *(log2 scale)*

```{r, warning=FALSE}
last.days(last.days.dat, 'death', params$last.days, countries, log2.flag = TRUE, per.100k.flag = TRUE)
```

### Data for Cases

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(last.days.dat, 'confirmed') %>% datatable
```

### Data for Deaths

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(last.days.dat, 'death') %>% datatable
```

# Rolling average of previous 4 days *(tracking over time)* {.tabset .tabset-fade .tabset-pills}

**Warning**: This visualization is not very intuitive, use it to see the trend over the last days.

## Cases

```{r}
last.week.cumulative(last.week.dat, 'confirmed', 4, countries, log2.flag = FALSE, per.100k.flag = FALSE)
```

## Deaths

```{r}
last.week.cumulative(last.week.dat, 'death', 4, countries, log2.flag = FALSE, per.100k.flag = FALSE)
```

## Cases *(per 100k population)*

```{r}
last.week.cumulative(last.week.dat, 'confirmed', 4, countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

## Deaths *(per 100k population)*

```{r}
last.week.cumulative(last.week.dat, 'death', 4, countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

## Data for Cases

```{r}
cumulative.last.days.data(last.week.dat, 'confirmed', countries, 4) %>% datatable()
```

## Data for Deaths

```{r}
cumulative.last.days.data(last.week.dat, 'death', countries, 4) %>% datatable()
```

# Cases vs. Death visualizations *(per 100k population)* {.tabset .tabset-fade .tabset-pills}

This visualization shows the number of cases/deaths per 100k of the population.

* **x axis**: Confirmed cases per 100k population
* **y axis**: Deaths per 100k population
* **size of data point**: Total population of the `r region(1, TRUE)`
* **label**: Rate of deaths per confirmed cases

## `r region(2, TRUE)`

```{r}
death.vs.cases.plot(death.vs.cases, state.filter = countries, expand = FALSE)
```

