# World  {.tabset .tabset-fade .tabset-pills}

Global data that sums all the `r region.code`-level confirmed cases and deaths.

## Totals

Total cumulative numbers.

```{r}
ommit.start(my.world$main, 'all', params$confirmed.start)
```

## log2 scale

Total cumulative numbers in log2 scale.

```{r, warning=FALSE}
ommit.start(my.world$main, 'all', params$confirmed.start, log2.flag = TRUE)
```

## New

New cases per day. *(warning: data for today is partial)*

```{r}
last.days(my.world$last.days, 'all', params$last.days, c(), log2.flag = FALSE, per.100k.flag = FALSE, new.flag = TRUE)
```

## Average of previous 4 days

Rolling average of previous 4 days. *(warning: data for today is partial)*

```{r}
last.week.cumulative(my.world$last.week, 'all', 4, c(), log2.flag = FALSE, per.100k.flag = FALSE)
```

## Average of previous 4 days *(log2 scale)*

*(warning: data for today is partial)*

```{r}
last.week.cumulative(my.world$last.week, 'all', 4, c(), log2.flag = TRUE, per.100k.flag = FALSE)
```


# World per 100k population  {.tabset .tabset-fade .tabset-pills}

## Totals

Total cumulative numbers per 100k population. *(warning: data for today is partial)*

```{r}
ommit.start(my.world$main, 'all', params$confirmed.start, per.100k.flag = TRUE)
```

## *log2 scale*

Total cumulative numbers in log2 scale per 100k population. *(warning: data for today is partial)*

```{r, warning=FALSE}
ommit.start(my.world$main, 'all', params$confirmed.start, per.100k.flag = TRUE, log2.flag = TRUE, digits = 2)
```

## New Cases

New cases per day per 100k population. *(warning: data for today is partial)*

```{r}
last.days(my.world$last.days, 'all', params$last.days, c(), log2.flag = FALSE, per.100k.flag = TRUE, new.flag = TRUE, digits = 4)
```

## Average of previous 4 days

Rolling average of previous 4 days per 100k population. *(warning: data for today is partial)*

```{r}
last.week.cumulative(my.world$last.week, 'all', 4, c(), log2.flag = FALSE, per.100k.flag = TRUE)
```

## Average of previous 4 days *(log2 scale)*

Rolling average of previous 4 days per 100k population. *(warning: data for today is partial)*

```{r}
last.week.cumulative(my.world$last.week, 'all', 4, c(), log2.flag = TRUE, per.100k.flag = TRUE)
```

