---
title: "Germany covid-19 analysis"
output: 
    html_document:
        toc: true
        dev: 'svg'
params:
    death.start: !r 5
    confirmed.start: !r 100
    duplicate.every: !r 2
    last.days: !r 12
---

# Preface

This is a personal analysis of the cases of covid-19.

I'm not from the Epidemiologist! My background is computer science and bioinformatics.

## Other covid-19 analysis

* [World](https://averissimo.github.io/covid19-analysis/)
* [Italy (by regione)](https://averissimo.github.io/covid19-analysis/italy.html)
* [Bavaria (in Germany)](https://averissimo.github.io/covid19-analysis/bayern.html)

```{r render.me, include=FALSE, eval=FALSE}
rmarkdown::render('germany.Rmd', quiet = TRUE)
```

```{r render.site, include=FALSE, eval=FALSE}
rmarkdown::render_site(quiet = TRUE)
rmarkdown::render_site()
rmarkdown::clean_site()
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, collapse = TRUE, fig.width = 10, fig.height = 8)
library(tidyverse)
library(dplyr)
library(viridis) # color pallete
library(ggrepel)
library(zoo)
library(glue)
library(DT)
library(anytime)
library(loose.rock)
library(scales)
library(futile.logger)
library(wbstats) # world bank data
library(reshape2)
library(eurostat)

devtools::load_all()

Sys.setlocale('LC_TIME', 'en_GB.UTF-8')
.Last.value <- flog.layout(layout.format('~m'))
```

# Data

Data was retrieved from *'RamiKrispin/coronavirus'* R data package which is updated daily in github.com.

It's not realtime data and data presented may have 1 or more days of difference from real-time data. Check the caption on the figures to check when was the last data point.


```{r buildCorona, echo=FALSE, message=FALSE, warning=FALSE}
eu.de <- download.de.data()

my.corona.original <- eu.de$data

source.by = eu.de$source
```

```{r eurostat}
populations <- download.eurostat.population('DE')

my.corona <- my.corona.original %>% left_join(populations, by = 'state')

my.corona <- my.corona %>% 
    ungroup %>%
    mutate(state.data = factor(state, levels = my.corona %>% filter(type == 'confirmed') %>% summarise(cases = max(cumul)) %>% arrange(cases) %>% pull(state)),
           state = state) %>%
    group_by(state)
```

```{r}
last.date          <- format(max(my.corona.original$date), '%Y/%m/%d')
last.date.string   <- paste0('Latest date from ', last.date, ' (source: ', source.by, ')')
region.code        <- 'State'
region.code.plural <- 'States'

countries                <- my.corona$state %>% unique
countries.always.include <- my.corona$state %>% unique
countries.extended       <- countries.always.include
```

```{r, include=FALSE, eval=FALSE}
#
#
#
#
#
#
#
#
#
#
#
#
#
#     BELOW should be always the same
#
#
#
#
#
#
#
#
#
#
#
#

```


```{r, collapse=TRUE}
flog.info('### Summary of data ###')
flog.info('')
flog.info('  %s: %d', region(2), my.corona.original %>% filter(cases > 0) %>% pull(state) %>% unique %>% length)
flog.info('')
flog.info('      Cases: %d', sum((my.corona.original %>% filter(type == 'confirmed'))$cases))
flog.info('     Deaths: %d', sum((my.corona.original %>% filter(type == 'death'))$cases))
flog.info('')
flog.info('     Latest: %s from %s', last.date, source.by)
```


## Main `r region(2)` in visualizations

The plots presented get very complex as more `r region(2, TRUE)` are added. I'm focusing mainly on Europe and comparing against Portugal and Germany *(where I'm from vs. where I live)*

```{r, collapse=TRUE}
flog.info('Looking only at the following %s: \n\n  %s' , region(2, TRUE), paste0(countries, collapse = ', '))
```

# Top 30 `r region(2)` {.tabset .tabset-fade .tabset-pills}

## Total Cases

```{r, message=FALSE}
top30(my.corona.original, 'confirmed')
```

## Total Deaths

```{r, message=FALSE}
top30(my.corona.original, 'death')
```

## Data Cases

```{r}
my.corona.original %>% top30.data('confirmed') %>% datatable
```

## Data Deaths

```{r}
my.corona.original %>% top30.data('death') %>% datatable
```

```{r}
my.corona.after.100 <- filter.after(my.corona, params$death.start, params$confirmed.start)
```

# Confirmed cases / Deaths {.tabset .tabset-fade .tabset-pills}

## Confirmed

```{r}
ommit.start(my.corona, 'confirmed', params$confirmed.start, filter.states = countries)
```

## Deaths

```{r}
ommit.start(my.corona, 'death', params$confirmed.start, filter.states = countries)
```

## Confirmed *(log2 scale)*

```{r}
ommit.start(my.corona, 'confirmed', params$confirmed.start, filter.states = countries, log2.flag = TRUE)
```

## Deaths *(log2 scale)*

```{r}
ommit.start(my.corona, 'death', params$confirmed.start, filter.states = countries, log2.flag = TRUE)
```

## Data for Confirmed

```{r}
ommit.data(my.corona, 'confirmed') %>% datatable
```

## Data for Deaths

```{r}
ommit.data(my.corona, 'confirmed') %>% datatable
```

# Confirmed cases / Deaths *(ommiting early start of epidemic)*

## Starting point after exceeding `r params$confirmed.start` cases or `r params$death.start` deaths {.tabset .tabset-fade .tabset-pills}

The visualizations presented below have been time-shifted to the day when each `r region(1, TRUE)` exceeded `r params$confirmed.start` cases or `r params$death.start` deaths *(depending on the visualization)*.

It is an arbitrary number, but allows to have a more accurate view of the progression as the first cases are quite sporadic and make the plots more complex.

**Important note**: All plots are in the logarithm scale *(base 2)* as it allows to better visualize when cases duplicate.

It is used because covid-19 cases grow very fast, and this way it becomes easier to read over time the number of cases. Always look at the y axis.

If you do not understand logarythm scale, I recommend this [video](https://www.youtube.com/watch?v=Kas0tIxDvrg).

### Cases  

```{r}
ommit.start(my.corona.after.100, 'confirmed', params$confirmed.start, filter.states = countries)
```

### Deaths

```{r}
ommit.start(my.corona.after.100, 'death', params$death.start, filter.states = countries)
```

### Cases *(log2 scale)*

```{r}
ommit.start(my.corona.after.100, 'confirmed', params$confirmer.start, filter.states = countries, log2.flag = TRUE)
```

### Deaths *(log2 scale)*

```{r}
ommit.start(my.corona.after.100, 'death', params$death.start, filter.states = countries, log2.flag = TRUE)
```

### Data for Cases

```{r}
my.corona.after.100.confirmed.dat <- ommit.data(my.corona.after.100, 'confirmed', countries)
my.corona.after.100.confirmed.dat %>% datatable
```

### Data for Deaths

```{r}
my.corona.after.100.death.dat <- ommit.data(my.corona.after.100, 'death', countries)
my.corona.after.100.death.dat %>% datatable()
```

## Starting point after exceeding `r params$confirmed.start` cases or `r params$death.start` deaths *(per 100k population)* {.tabset .tabset-fade .tabset-pills}

Showing population for plotted `r region(2, TRUE)` This becomes relevant for the next plots.

### Cases

```{r}
ommit.start(my.corona.after.100, 'confirmed', params$confirmed.start, filter.states = countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

### Deaths 

```{r}
ommit.start(my.corona.after.100, 'death', params$confirmer.start, filter.states = countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

### Cases *(log2 scale)*

```{r}
ommit.start(my.corona.after.100, 'confirmed', params$confirmer.start, filter.states = countries, log2.flag = TRUE, per.100k.flag = TRUE)
```

### Deaths *(log2 scale)*

```{r}
ommit.start(my.corona.after.100, 'death', params$confirmer.start, filter.states = countries, log2.flag = TRUE, per.100k.flag = TRUE)
```

### Data for Cases

```{r}
my.corona.after.100.confirmed.dat %>% datatable
```

### Data for Deaths

```{r}
my.corona.after.100.death.dat %>% datatable
```

# Last `r params$last.days` days {.tabset .tabset-fade .tabset-pills}

Instead of showing the evolution of the cases/deaths since a `r region(1, TRUE)` exceeded `r params$confirmed.start`/`r params$death.start` cases, it insteads shows the last 12 days.

The first data point is the number of new cases that day.

## Cases 

```{r}
my.corona.last.days <- filter.last.days(my.corona, params$last.days)
last.days(my.corona.last.days, 'confirmed', params$last.days, countries, log2.flag = FALSE, per.100k.flag = FALSE)
```

## Deaths

```{r, warning=FALSE}
last.days(my.corona.last.days, 'death', params$last.days, countries, log2.flag = FALSE, per.100k.flag = FALSE)
```

## Cases *(log2 scale)*

```{r, message=FALSE, warning=FALSE}
last.days(my.corona.last.days, 'confirmed', params$last.days, countries, log2.flag = TRUE, per.100k.flag = FALSE)
```

## Deaths

```{r, warning=FALSE}
last.days(my.corona.last.days, 'death', params$last.days, countries, log2.flag = TRUE, per.100k.flag = FALSE)
```

## Data for Cases

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(my.corona.last.days, 'confirmed', countries) %>% datatable
```

## Data for Deaths

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(my.corona.last.days, 'death', countries) %>% datatable
```

# Last `r params$last.days` days *(per 100k population)* {.tabset .tabset-fade .tabset-pills}

Instead of showing the evolution of the cases/deaths since a `r region(1, TRUE)` exceeded `r params$confirmed.start`/`r params$death.start` cases, it insteads shows the last 12 days.

The first data point is the number of new cases that day.

## Cases 

```{r}
my.corona.last.days <- filter.last.days(my.corona, params$last.days)
last.days(my.corona.last.days, 'confirmed', params$last.days, countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

## Deaths

```{r, warning=FALSE}
last.days(my.corona.last.days, 'death', params$last.days, countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

## Cases *(log2 scale)*

```{r, message=FALSE, warning=FALSE}
last.days(my.corona.last.days, 'confirmed', params$last.days, countries, log2.flag = TRUE, per.100k.flag = TRUE)
```

## Deaths *(log2 scale)*

```{r, warning=FALSE}
last.days(my.corona.last.days, 'death', params$last.days, countries, log2.flag = TRUE, per.100k.flag = TRUE)
```

## Data for Cases

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(my.corona.last.days, 'confirmed', countries) %>% datatable
```

## Data for Deaths

**warning**: sub-total is only for last 12 days

```{r}
last.days.data(my.corona.last.days, 'death', countries) %>% datatable
```


# Sum of previous 4 days *(tracking over time)* {.tabset .tabset-fade .tabset-pills}

**Warning**: This visualization is not very intuitive, use it to see the trend over the last days.

```{r}
last.week <- filter.last.days.cumulative(my.corona, 4)
```

## Cases

```{r}
last.week.cumulative(last.week, 'confirmed', 4, countries, log2.flag = FALSE, per.100k.flag = FALSE)
```

## Deaths

```{r}
last.week.cumulative(last.week, 'death', 4, countries, log2.flag = FALSE, per.100k.flag = FALSE)
```

## Cases *(per 100k population)*

```{r}
last.week.cumulative(last.week, 'confirmed', 4, countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

## Deaths *(per 100k population)*

```{r}
last.week.cumulative(last.week, 'death', 4, countries, log2.flag = FALSE, per.100k.flag = TRUE)
```

## Data for Cases in last 4 days

```{r}
cumulative.last.days.data(last.week, 'confirmed', countries, 4) %>% datatable()
```

## Data for Deaths in last 4 days

```{r}
cumulative.last.days.data(last.week, 'death', countries, 4) %>% datatable()
```

# Cases vs. Death visualizations *(per 100k population)* {.tabset .tabset-fade .tabset-pills}

This visualization shows the number of cases/deaths per 100k of the population.

* **x axis**: Confirmed cases per 100k population
* **y axis**: Deaths per 100k population
* **size of data point**: Total population of the `r region(1, TRUE)`
* **label**: Rate of deaths per confirmed cases

```{r}
death.vs.cases <- filter.death.vs.cases(last.week) %>% filter(state %in% c(countries.extended, countries.always.include) %>% unique)
```

## All states

```{r}
death.vs.cases.plot(death.vs.cases, always.include = countries.always.include)
```

## Data

```{r}
death.vs.cases.data(death.vs.cases) %>% datatable()
```

## Number of Confirmed Cases

```{r}
my.corona.original %>% filter(state %in% countries.extended) %>% cases.plot('confirmed')
```

## Number of Deaths

```{r}
my.corona.original %>% filter(state %in% countries.extended) %>% cases.plot('death')
```
