---
title: "Personal analysis of covid-19 cases"
output: 
    html_document:
        toc: true
        dev: 'svg'
params:
    death.start: !r 10
    confirmed.start: !r 100
    duplicate.every: !r 2
    last.days: !r 12
---

# Preface

This is a personal analysis of the cases of covid-19.

I'm not from the Epidemiology field. I'm currently working on modeling cancer datasets for my PhD thesis.

```{r render.me, include=FALSE, eval=FALSE}
rmarkdown::render('Analysis.Rmd', quiet = TRUE)
```

```{r render.site, include=FALSE, eval=FALSE}
rmarkdown::render_site(quiet = TRUE)
rmarkdown::clean_site()
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 8)
library(tidyverse)
library(dplyr)
library(viridis) # color pallete
library(ggrepel)
library(zoo)
library(glue)
library(DT)
library(scales)
library(futile.logger)
library(wbstats) # world bank data

.Last.value <- flog.layout(layout.format('~m'))
```

# Data

Data was retrieved from *'RamiKrispin/coronavirus'* R data package which is updated daily in github.com.

It's not realtime data and data presented may have 1 or more days of difference from real-time data. Check the caption on the figures to check when was the last data point.

```{r reinstall.data.package, echo=FALSE, message=FALSE, warning=FALSE}
BiocManager::install('RamiKrispin/coronavirus', ask = FALSE)

library(coronavirus)

data("coronavirus")
```

```{r buildCorona, include=FALSE}
my.corona.original <- coronavirus %>% as_tibble() %>%
    mutate(country = Country.Region) %>%
    group_by(country, type, date) %>%
    summarise(cases = sum(cases)) %>%
    arrange(date) %>%
    mutate(cumul = cumsum(cases))
```

```{r summaryCorona, echo=FALSE, collapse=TRUE}
flog.info('### Summary of data ###')
flog.info('')
flog.info('  Countries: %d', my.corona.original %>% filter(cases > 0) %>% pull(country) %>% unique %>% length)
flog.info('')
flog.info('      Cases: %d', sum((my.corona.original %>% filter(type == 'confirmed'))$cases))
flog.info('  Recovered: %d', sum((my.corona.original %>% filter(type == 'recovered'))$cases))
flog.info('     Deaths: %d', sum((my.corona.original %>% filter(type == 'death'))$cases))
flog.info('')

last.date <- format(max(my.corona.original$date), '%Y/%m/%d')
last.date.string <- paste0('Latest date from ', last.date, ' (source: RamiKrispin/coronavirus r package)')

flog.info('     Latest: %s', last.date)
```

## Countries in visualizations

The plots presented get very complex as more countries are added. I'm focusing mainly on Europe and comparing against Portugal and Germany *(where I'm from vs. where I live)*

```{r, echo=FALSE, collapse=TRUE}
countries <- c('Italy', 'Spain','Germany', 'China', 'Portugal', 'US', 'France', 'Switzerland')

countries <- c('Italy', 'Spain','Germany', 'China', 'Portugal', 'US', 'France', 'Switzerland')

flog.info('Looking only at the following countries: \n\n  %s' , paste0(countries, collapse = ', '))
```

# Top 30 countries {.tabset .tabset-fade .tabset-pills}

## Total Confirmed cases

```{r top30confirmed, echo=FALSE}
my.corona.original %>%
    filter(type == 'confirmed') %>%
    group_by(country, type) %>%
    summarise(cases = sum(cases)) %>%
    arrange(cases) %>%
    ungroup() %>%
    mutate(country = paste0(country, ' (', cases, ')')) %>%
    mutate(country = factor(country, levels = unique(.$country))) %>%
    top_n(30, cases) %>%
    ggplot() +
        geom_bar(aes(country, cases, fill = country), stat = 'identity') +
        # scale_y_continuous(trans = 'log10') +
        coord_flip() +
        labs(x = 'Confirmed Cases', 
             y = '', 
             title = 'Confirmed Cases by Country' , 
             susubtitle = 'only showing top 30',
             caption = last.date.string) + 
        theme_minimal() + 
        theme(legend.position = 'none')
```

## Total Deaths

```{r top30deaths, echo=FALSE}
my.corona.original %>%
    filter(type == 'death') %>%
    group_by(country, type) %>%
    summarise(cases = sum(cases)) %>%
    arrange(cases) %>%
    ungroup() %>%
    mutate(country = paste0(country, ' (', cases, ')')) %>%
    mutate(country = factor(country, levels = unique(.$country))) %>%
    filter(cases > 0) %>%
    top_n(30, cases) %>%
    ggplot() +
        geom_bar(aes(country, cases, fill = country), stat = 'identity') +
        # scale_y_continuous(trans = 'log10') +
        coord_flip() +
        labs(x = 'Deaths', 
             y = '', 
             title = 'Deaths by Country' , 
             subtitle = 'only showing top 30',
             caption = last.date.string) + 
        theme_minimal() + 
        theme(legend.position = 'none')
```

## Data confirmed

```{r data.confirmed}
my.corona.original %>%
    filter(type == 'confirmed') %>%
    group_by(country, type) %>%
    summarise(cases = sum(cases)) %>%
    arrange(cases) %>%
    ungroup() %>%
    filter(cases > 0) %>%
    arrange(-cases) %>%
    mutate(cases = format(cases, big.mark = ',')) %>%
    select(Country = country, 'Cases Confirmed' = cases) %>%
    datatable()
```

## Data death

```{r data.death}
my.corona.original %>%
    filter(type == 'death') %>%
    group_by(country, type) %>%
    summarise(cases = sum(cases)) %>%
    arrange(cases) %>%
    ungroup() %>%
    filter(cases > 0) %>%
    arrange(-cases) %>%
    mutate(cases = format(cases, big.mark = ',')) %>%
    select(Country = country, 'Cases Death' = cases) %>%
    datatable()
```


```{r worldbank, echo=FALSE, collapse=TRUE}
### Add census data from World Bank
#
# Population census data from World Bank
#

# Get data from world bank

populations <- wb(indicator = "SP.POP.TOTL") %>% 
    group_by(iso3c) %>%
    filter(date == max(date)) %>%
    select(population.year = date, population = value, country, iso3c) %>%
    arrange(country)

# Change some country name to match covid-19 data

convert.names <- list('United States' = 'US',
     'Czech Republic' = 'Czechia',
     'Russian Federation' = 'Russia',
     'Slovak Republic' = 'Slovakia',
     'Korea, Rep.' = 'Korea, South')

for (ix.name in names(convert.names)) {
    populations <- populations %>%
        arrange(country) %>%
        ungroup %>%
        mutate(country = replace(country, country == ix.name, convert.names[[ix.name]]))    
}




# info

#flog.info('Countries with names different in world bank: \n\n  %s',
#          paste0(unique(populations$country)[!unique(populations$country) %in% levels(my.corona.original$country)], 
#                 collapse = ', '))

# Join the two together, keeping all of covid-19 original data

my.corona <- my.corona.original %>% 
    left_join(populations, by = 'country')

my.corona <- mutate(my.corona, 
                    country.data = factor(country, levels = my.corona %>% filter(type == 'confirmed') %>% summarise(cases = max(cumul)) %>% arrange(cases) %>% pull(country))) %>%
    select(-iso3c)
```

```{r, echo=FALSE}
my.corona.after.100 <- my.corona %>%
    #
    filter((type == 'death' & cumul >= params$death.start ) | 
           (type == 'confirmed' & cumul >= params$confirmed.start) | 
           (type == 'recovered' & cumul >= 0)) %>%
    mutate(days.after.100 = as.numeric(difftime(date, min(date), units = 'days')))
```

# Confirmed cases / Deaths *(ommiting start of epidemic)*

## Starting point after exceeding 100 cases or 10 deaths {.tabset .tabset-fade .tabset-pills}

The visualizations presented below have been time-shifted to the day when each country exceeded 100 cases or 10 deaths *(depending on the visualization)*.

It is an arbitrary number, but allows to have a more accurate view of the progression as the first cases are quite sporadic and make the plots more complex.

**Important note**: All plots are in the logarithm scale *(base 2)* as it allows to better visualize when cases duplicate.

It is used because covid-19 cases grow very fast, and this way it becomes easier to read over time the number of cases. Always look at the y axis.

If you do not understand logarythm scale, I recommend this [video](https://www.youtube.com/watch?v=Kas0tIxDvrg).

### Confirmed cases  

```{r corfirmed.cases.after.100, echo=FALSE}
my.corona.after.100 %>%
    filter(country %in% countries) %>%
    filter(type == 'confirmed') %>%
    ungroup(country) %>%
    group_by(country) %>%
    arrange(-cases) %>% {
        tmp <- summarise(., cases = max(cumul))
        tmp <- arrange(tmp, cases)
        tmp <- mutate(tmp, country.data = paste0(country, ' (', cases, ')'))
        
        mutate(., country.data = factor(paste0(country, ' (', max(cumul), ')'),
                                        levels = rev(pull(tmp, country.data))))
        } %>%
    mutate(label = if_else(cumul == max(cumul), as.character(country.data), NA_character_)) %>%
    ungroup %>%
    ggplot(aes(x = days.after.100, y = cumul, color = country.data)) + 
        geom_line(size = 1.2) +
        geom_label_repel(aes(label = label,
                             fill = country.data), 
                         na.rm = TRUE,
                         color = 'white',
                         size = 3.5,
                         segment.alpha = .4,
                         segment.colour = 'black') +
        labs(y = 'Confirmed cases',
             x = 'Number of days since ≥ 100 confirmed cases',
             title = 'Confirmed cases over time',
             subtitle = 'Only showing after each country had more than 100 cases',
             caption = last.date.string) +
        scale_y_continuous('Confirmed cases (log2 scale)', trans = 'log2') + 
        scale_color_viridis(discrete = TRUE, end = .85, option = 'A') +
        scale_fill_viridis(discrete = TRUE, end = .85, option = 'A') +
        theme_minimal() +
        theme(legend.position = 'none', legend.title = element_blank())
```

### Patient deaths

```{r death.after.100, echo=FALSE}
my.corona.after.100 %>%
    filter(country %in% countries) %>%
    filter(type == 'death') %>%
    group_by(country) %>% {
        tmp <- summarise(., cases = max(cumul))
        tmp <- arrange(tmp, cases)
        tmp <- mutate(tmp, country.data = paste0(country, ' (', cases, ')'))
        
        mutate(., country.data = factor(paste0(country, ' (', max(cumul), ')'),
                                        levels = rev(pull(tmp, country.data))))
        } %>%
    mutate(label = if_else(cumul == max(cumul), as.character(country.data), NA_character_)) %>%
    ungroup %>%
    ggplot(aes(x = days.after.100, y = cumul, color = country.data)) + 
        # geom_point(aes(date, cumul, color = country)) +
        geom_line(size = 1.2) +
        geom_label_repel(aes(label = label,
                             fill = country.data), 
                             na.rm = TRUE,
                             color = 'white',
                             size = 3.5,
                             segment.alpha = .4,
                             segment.colour = 'black') +
        labs(y = 'Deaths',
             x = 'Number of days since ≥ 10 deaths',
             title = 'Deaths over Time',
             subtitle = 'Only showing after each country had more than 10 deaths',
             caption = last.date.string) +
        scale_y_continuous('Deaths (log2 scale)', trans = 'log2') + 
        scale_color_viridis(discrete = TRUE, end = .85, option = 'A') +
        scale_fill_viridis(discrete = TRUE, end = .85, option = 'A') +
        theme_minimal() +
        theme(legend.position = 'none', legend.title = element_blank())
```

## Starting point after exceeding 100 cases or 10 deaths *(per 100k population)* {.tabset .tabset-fade .tabset-pills}

Showing population for plotted countries. This becomes relevant for the next plots.

### Confirmed cases

```{r confirmed.after.100.per100k, echo=FALSE}
my.corona.after.100 %>%
    filter(country %in% countries) %>%
    filter(type %in% c('confirmed')) %>%
    mutate(cumul.per.100k = cumul / population * 100000) %>%
    group_by(country) %>% {
        tmp <- summarise(., cases = round(max(cumul.per.100k), digits = 2))
        tmp <- arrange(tmp, cases)
        tmp <- mutate(tmp, country.data = paste0(country, ' (', cases, ')'))
        
        mutate(., country.data = factor(paste0(country, ' (', round(max(cumul.per.100k), digits = 2), ')'),
                                        levels = rev(pull(tmp, country.data))))
        } %>%
    mutate(label = if_else(cumul.per.100k == max(cumul.per.100k), as.character(country.data), NA_character_)) %>%
    ungroup %>%
    ggplot(aes(x = days.after.100, y = cumul.per.100k, color = country.data)) + 
        geom_line(size = 1.2) +
        geom_label_repel(aes(label = label,
                                 fill = country.data), 
                             na.rm = TRUE,
                             color = 'white',
                             size = 3.5,
                             segment.alpha = .4,
                             segment.colour = 'black') +
        labs(x = 'Number of days since ≥ 100 cases',
             y = 'Confirmed cases -- per 100k population',
             title = 'Confirmed cases per 100k population',
             subtitle = 'Only showing after each country had more than 100 cases',
             caption = last.date.string) + 
        scale_y_continuous('Confirmed cases -- per 100k population (log2 scale)', trans = 'log2') +
        scale_color_viridis(discrete = TRUE, end = .85, option = 'C') +
        scale_fill_viridis(discrete = TRUE, end = .85, option = 'C') +
        theme_minimal() +
        theme(legend.position = 'none', legend.title = element_blank())
```

### Deaths 

```{r death.after.100.per.100k, echo=FALSE}
my.corona.after.100 %>%
    filter(country %in% countries) %>%
    filter(type %in% c('death')) %>%
    mutate(cumul = cumul / population * 100000) %>%
    group_by(country) %>% {
        tmp <- summarise(., cases = round(max(cumul), digits = 2))
        tmp <- arrange(tmp, cases)
        tmp <- mutate(tmp, country.data = paste0(country, ' (', cases, ')'))
        
        mutate(., country.data = factor(paste0(country, ' (', round(max(cumul), digits = 2), ')'),
                                        levels = rev(pull(tmp, country.data))))
        } %>%
    mutate(label = if_else(cumul == max(cumul), as.character(country.data), NA_character_)) %>%
    ungroup %>%
    ggplot(aes(x = days.after.100, y = cumul, color = country.data)) + 
        geom_line(size = 1.2) +
        geom_label_repel(aes(label = label,
                                 fill = country.data), 
                             na.rm = TRUE,
                             color = 'white',
                             size = 3.5,
                             segment.alpha = .4,
                             segment.colour = 'black') +
        labs(x = 'Number of days since ≥ 10 deaths',
             y = 'Deaths -- per 100k population',
             title = 'Deaths per 100k population',
             subtitle = 'Only showing after each country had more than 10 deaths',
             caption = last.date.string) + 
        scale_y_continuous('Deaths -- per 100k population (log2 scale)', trans = 'log2') + 
        scale_color_viridis(discrete = TRUE, end = .85, option = 'C') +
        scale_fill_viridis(discrete = TRUE, end = .85, option = 'C') +
        theme_minimal() +
        theme(legend.position = 'none', legend.title = element_blank())
```

### Population data

```{r population.data, echo=FALSE}
my.corona %>% 
    filter(country %in% countries) %>% 
    ungroup %>% 
    select(country, population, population.year) %>% 
    distinct(country, .keep_all = TRUE) %>%
    arrange(population) %>%
    mutate(population = format(population, big.mark = ',', trim = FALSE)) %>%
    datatable()
```

# Last `r params$last.days` days {.tabset .tabset-fade .tabset-pills}

Instead of showing the evolution of the cases/deaths since a country exceeded 100/10 cases, it insteads shows the last 12 days.

The first data point is the number of new cases that day.

## Confirmed cases 

```{r confirmed.last12.days, echo=FALSE}
my.corona.after.100 %>%
    filter(country %in% countries) %>%
    filter(type == 'confirmed') %>%
    arrange(country) %>%
    top_n(params$last.days, days.after.100) %>%
    mutate(days.before.now = days.after.100 - max(days.after.100)) %>%
    mutate(cumul = cumsum(cases)) %>%
    group_by(country) %>% {
        tmp <- summarise(., cases = max(cumul))
        tmp <- arrange(tmp, cases)
        tmp <- mutate(tmp, country.data = paste0(country, ' (', cases, ')'))
        
        mutate(., country.data = factor(paste0(country, ' (', max(cumul), ')'),
                                        levels = rev(pull(tmp, country.data))))
        } %>%
    mutate(label = if_else(cumul == max(cumul), as.character(country.data), NA_character_)) %>%
    ungroup %>%
    ggplot(aes(x = days.before.now, y = cumul, color = country.data)) + 
        geom_line(size = 1.2) +
        geom_label_repel(aes(label = label,
                                 fill = country.data), 
                             na.rm = TRUE,
                             color = 'white',
                             size = 3.5,
                             segment.colour = 'black',
                             segment.alpha = .4) +
        labs(x = paste0('Last ', params$last.days, ' days'),
             y = 'Confirmed cases',
             title = paste0('Confirmend Cases in the last ', params$last.days, ' days'),
             caption = last.date.string) + 
        scale_y_continuous('Confirmed cases (log2 scale)', trans = 'log2') + 
        scale_color_viridis(discrete = TRUE, end = .85, option = 'A') +
        scale_fill_viridis(discrete = TRUE, end = .85, option = 'A') +
        #
        theme_minimal() +
        theme(legend.position = 'none', legend.title = element_blank())
```

## Deaths

```{r death.last12.days, echo=FALSE}
my.corona.after.100 %>%
    filter(country %in% countries) %>%
    filter(type == 'death') %>%
    arrange(country) %>%
    top_n(params$last.days, days.after.100) %>%
    mutate(days.before.now = days.after.100 - max(days.after.100)) %>%
    mutate(cumul = cumsum(cases)) %>%
    group_by(country) %>% {
        tmp <- summarise(., cases = max(cumul))
        tmp <- arrange(tmp, cases)
        tmp <- mutate(tmp, country.data = paste0(country, ' (', cases, ')'))
        
        mutate(., country.data = factor(paste0(country, ' (', max(cumul), ')'),
                                        levels = rev(pull(tmp, country.data))))
        } %>%
    mutate(label = if_else(cumul == max(cumul), as.character(country.data), NA_character_)) %>%
    ungroup %>%
    ggplot(aes(x = days.before.now, y = cumul, color = country.data)) + 
        geom_line(size = 1.2) +
        geom_label_repel(aes(label = label,
                                 fill = country.data), 
                             na.rm = TRUE,
                             color = 'white',
                             size = 3.5,
                             segment.alpha = .4,
                             segment.colour = 'black') +
        labs(x = paste0('Last ', params$last.days, ' days'),
             y = 'Deaths',
             title = paste0('Deaths in the last ', params$last.days, ' days'),
             caption = last.date.string) + 
        scale_y_continuous('Deaths (log2 scale)', trans = 'log2') + 
        scale_color_viridis(discrete = TRUE, end = .85, option = 'B') +
        scale_fill_viridis(discrete = TRUE, end = .85, option = 'B') +
        theme_minimal() +
        theme(legend.position = 'none', legend.title = element_blank())
```

## Data for confirmed cases

```{r last.12.days.confirmed.data, echo=FALSE}
my.corona.after.100 %>%
    filter(country %in% countries) %>%
    filter(type == 'confirmed') %>%
    arrange(country) %>%
    top_n(params$last.days, days.after.100) %>%
    mutate(days.before.now = days.after.100 - max(days.after.100)) %>%
    mutate(cumul = cumsum(cases)) %>%
    ungroup() %>%
    mutate(Country = country,
           Date = format(date, '%B %d'),
           'Confirmed cases' = cases,
           'Total cases' = cumul) %>%
    select(Country, Date, 'Confirmed cases', 'Total cases') %>%
    datatable()
```

## Data for Deaths

```{r last.12.days.death.data, echo=FALSE}
my.corona.after.100 %>%
    filter(country %in% countries) %>%
    filter(type == 'death') %>%
    arrange(country) %>%
    top_n(params$last.days, days.after.100) %>%
    mutate(days.before.now = days.after.100 - max(days.after.100)) %>%
    mutate(cumul = cumsum(cases)) %>%
    ungroup() %>%
    mutate(Country = country,
           Date = format(date, '%B %d'),
           'Deaths' = cases,
           'Total deaths' = cumul) %>%
    select(Country, Date, 'Deaths', 'Total deaths') %>%
    datatable()
```

# Cumulative cases of previous 4 days *(how it tracks over time)* {.tabset .tabset-fade .tabset-pills}

**Warning**: This visualization is not very intuitive, use it to see the trend over the last days.

```{r countries.more, echo=FALSE}
countries.eu <- c('Italy', 'Spain','Germany', 'Portugal', 'France',  
                  'United Kingdom', 'Poland', 'Greece', 'Croatia', 'Slovenia', 
                  'Hungary', 'Slovakia', 'Ireland', 'Luxembourg', 'Finland', 
                  'Sweden', 'Denmark', 'Romania', 'Netherlands', 'Belgium', 
                  'Czechia', 'Austria', 'Bulgaria', 'Lithuania', 'Latvia', 
                  'Malta', 'Estonia', 'Cyprus')

countries.cplp <- c('Angola', 'Mozambique', 'Cabo Verde', 'Timor-Leste', 
                    'Brazil'
                    , 'Macao SAR, China', 'Guinea-Bissau', 'Sao Tome and Principe' # not in corona data
                    )

countries.europe.non.eu <- c('Norway', 'Switzerland', 'Ukraine', 'Serbia', 'Russia', 
                             'Albania', 'Iceland', 'Turkey', 'Belarus')

countries.non.europe <- c('US', 'China', 'Canada', 
                          'Australia', 'New Zealand', 
                          'Korea, South', 'Japan', 'Argentina', 'Chile', 'Thailand')

countries.always.include <- c('Portugal', 'Iran', 'Germany', 'Spain', 'Italy', 'Korea, South')

countries.extended <- c(countries.eu, countries.europe.non.eu, countries.cplp, countries.non.europe) %>%
    unique

# countries.extended[!countries.extended %in% my.corona$country]
# countries.extended[!countries.extended %in% populations$country]

# my.corona$country %>% unique %>% sort
# populations$country %>% unique %>% sort

last.week.tmp <- my.corona.original %>% 
    filter(country %in% countries.extended) %>%
    # filter(country == 'Portugal') %>%
    group_by(country, type) %>%
    arrange(desc(date)) %>%
    ungroup() %>%
    select(country, date, type, cases)

last.week <- 
    # Join confirmed and death cases
    left_join(last.week.tmp %>% filter(type == 'confirmed') %>% select(-type),
                        last.week.tmp %>% filter(type == 'death') %>% select(-type),
                        by = c('country', 'date'),
                        suffix = c('.confirmed', '.death')) %>%
    #
    # Replace any NA from join
    mutate(cases.confirmed = replace(cases.confirmed, is.na(cases.confirmed), 0),
           cases.death = replace(cases.death, is.na(cases.death), 0)) %>%
    #    
    # Calculate cumulative sum over date
    arrange(date) %>%
    group_by(country) %>%
    #
    # calculate cumulative cases (i.e. all cases reported up to that day)
    mutate(cumul.death = cumsum(cases.death),
           cumul.confirmed = cumsum(cases.confirmed)) %>% 
    arrange(desc(date)) %>%
    mutate(last.week.cases.confirmed = rollapply(cases.confirmed, 4, sum, align = 'left', fill = c(0,0,0), partial = TRUE)) %>%
    mutate(last.week.cases.death = rollapply(cases.death, 4, sum, align = 'left', fill = c(0,0,0), partial = TRUE)) %>%
    #
    #
    left_join(populations, by = 'country') %>%
    group_by(country) %>%
    arrange(date) %>%
    #
    mutate(deaths.per.100k = last.week.cases.death / population * 100000,
           confirmed.per.100k = last.week.cases.confirmed / population * 100000,
           ratio.deaths.per.confirmed = last.week.cases.death / last.week.cases.confirmed) 
```

## Confirmed cases

```{r confirmed.lastweek.cumulative, echo=FALSE}
p <- last.week %>% 
    filter(country %in% countries) %>%
    group_by(country) %>% {
        tmp <- summarise(., cases = last(last.week.cases.confirmed))
        tmp <- arrange(tmp, cases)
        tmp <- mutate(tmp, country.data = paste0(country, ' (', cases, ')'))
        
        mutate(., country.data = factor(paste0(country, ' (', last(last.week.cases.confirmed), ')'),
                                        levels = rev(pull(tmp, country.data))))
    } %>%
    mutate(label = if_else(date == max(date), as.character(country.data), NA_character_)) %>%
    #
    ggplot(aes(x = date, y = last.week.cases.confirmed, color = country.data)) +
        geom_line(size = 1.2) +
        geom_point(size = 1.2) +
                geom_label_repel(aes(label = label,
                                     fill = country.data), 
                                 na.rm = TRUE,
                                 color = 'white',
                                 size = 3.5,
                                 segment.alpha = .4,
                                 segment.colour = 'black') +
        labs(title = 'WARNING:: Each point is a sum from previous 4 days',
             y = 'Number of cases confirmed in previous 4 days',
             x = 'Date',
             caption = last.date.string) +
        scale_color_viridis(discrete = TRUE, end = .85, option = 'A') +
        scale_fill_viridis(discrete = TRUE, end = .85, option = 'A') +
        theme_minimal() +
        theme(legend.position = 'none')
p
# p + scale_y_continuous('Number of cases confirmed in previous 4 days (log2 scale)', trans = 'log2')
```

## Deaths

```{r death.lastweek.cumulative, echo=FALSE}
p <- last.week %>% 
    filter(country %in% countries) %>%
    group_by(country) %>% {
        tmp <- summarise(., cases = last(last.week.cases.death))
        tmp <- arrange(tmp, cases)
        tmp <- mutate(tmp, country.data = paste0(country, ' (', cases, ')'))
        
        mutate(., country.data = factor(paste0(country, ' (', last(last.week.cases.death), ')'),
                                        levels = rev(pull(tmp, country.data))))
    } %>%
    mutate(label = if_else(date == max(date), as.character(country.data), NA_character_)) %>%
    #
    ggplot(aes(x = date, y = last.week.cases.death, color = country.data)) +
        geom_line(size = 1.2) +
                geom_label_repel(aes(label = label,
                                     fill = country.data), 
                                 na.rm = TRUE,
                                 color = 'white',
                                 size = 3.5,
                                 segment.alpha = .4,
                                 segment.colour = 'black') +
        labs(title = 'WARNING:: Each point is a sum from previous 4 days',
             y = 'Number of deaths in previous 4 days',
             x = 'Date',
             caption = last.date.string) +
        scale_color_viridis(discrete = TRUE, end = .85, option = 'A') +
        scale_fill_viridis(discrete = TRUE, end = .85, option = 'A') +
        theme_minimal() +
        theme(legend.position = 'none')
p
# p + scale_y_continuous('Number of cases confirmed in previous 4 days (log2 scale)', trans = 'log2')
```

# Confirmed cases vs. Death visualizations *(per 100k population)* {.tabset .tabset-fade .tabset-pills}

This visualization shows the number of cases/deaths per 100k of the population.

* **x axis**: Confirmed cases per 100k population
* **y axis**: Deaths per 100k population
* **size of data point**: Total population if the country
* **label**: Rate of deaths per confirmed cases

*note*: It always shows Portugal, Spain, Italy, South Korea, Iran and Germany.

```{r death.vs.cases.function.and.data, echo=FALSE}
death.vs.cases <- last.week %>%
    group_by(country, population) %>%
    summarise(ratio = max(cumul.death) / max(cumul.confirmed),
              cases.death = max(cumul.death / population * 100000), 
              cases.confirmed = max(cumul.confirmed / population * 100000)) %>%
    filter(cases.death > 0)

plot.death.vs.cases <- function(country.filter, always.include = countries.always.include) {
    return(death.vs.cases %>%
        filter(country %in% (c(always.include, country.filter) %>% unique)) %>%
        ggplot(aes(x = cases.confirmed, y = cases.death, color = country)) +
            geom_point(aes(size = population), alpha = .4) +
            geom_label_repel(aes(label = paste0(country, ' (', percent(ratio, accuracy = .1), ')'),
                                 fill = country),
                                 na.rm = TRUE,
                                 alpha = .8,
                                 color = 'white',
                                 size = 3,
                                 segment.alpha = .4,
                                 segment.colour = 'black',
                                 force = 2) +
            expand_limits(x = ceiling(max(death.vs.cases %>% pull(cases.confirmed))),
                          y = ceiling(max(death.vs.cases %>% pull(cases.death)))) +
            labs(x = 'Confirmed Cases per 100k population', 
                 y = 'Deaths per 100k population', 
                 title = 'Deaths vs. Cases per 100k population', 
                 subtitle = 'percentage is the death rate per confirmed cases and size represents size of country by population',
                 caption = last.date.string) +
            theme_minimal() +
            theme(legend.position = 'none'))
}
```

## EU

```{r eu, echo=FALSE}
plot.death.vs.cases(countries.eu)
```

## Non-EU Europe

```{r non.eu.europe, echo=FALSE}
plot.death.vs.cases(countries.europe.non.eu)
```

## Non-Europe

```{r non.europe, echo=FALSE}
plot.death.vs.cases(countries.non.europe)
```

## CPLP

```{r cplp, echo=FALSE}
plot.death.vs.cases(countries.cplp)
```

## Data for Confirmed Cases

```{r death.vs.cases.data.confirmed, echo=FALSE}
my.corona.original %>%
    filter(country %in% countries.extended) %>%
    filter(type == 'confirmed') %>%
    group_by(country, type) %>%
    summarise(cases = sum(cases)) %>%
    arrange(cases) %>%
    ungroup() %>%
    mutate(country = paste0(country, ' (', cases, ')')) %>%
    mutate(country = factor(country, levels = unique(.$country))) %>%
    ggplot() +
        geom_bar(aes(country, cases, fill = country), stat = 'identity') +
        # scale_y_continuous(trans = 'log10') +
        coord_flip() +
        labs(x = 'Confirmed Cases', 
             y = '', 
             title = 'Confirmed Cases by Country' ,
             caption = last.date.string) + 
        theme_minimal() + 
        theme(legend.position = 'none')
```

## Data for Deaths

```{r death.vs.cases.data.death, echo=FALSE}
my.corona.original %>%
    filter(country %in% countries.extended) %>%
    filter(type == 'death') %>%
    group_by(country, type) %>%
    summarise(cases = sum(cases)) %>%
    arrange(cases) %>%
    ungroup() %>%
    mutate(country = paste0(country, ' (', cases, ')')) %>%
    mutate(country = factor(country, levels = unique(.$country))) %>%
    ggplot() +
        geom_bar(aes(country, cases, fill = country), stat = 'identity') +
        # scale_y_continuous(trans = 'log10') +
        coord_flip() +
        labs(x = 'Deaths', 
             y = '', 
             title = 'Deaths by Country' , 
             caption = last.date.string) + 
        theme_minimal() + 
        theme(legend.position = 'none')
```

# Hospital beds vs. Cases *(per 100k population)* {.tabset .tabset-fade .tabset-pills}

This visualization shows the number of cases/deaths per 100k of the population.

* **x axis**: Confirmed cases per 100k population
* **y axis**: Hospilta beds per 100k population
* **size of data point**: Ratio of Confirmed cases / Hospital beds
* **label**: Country (<Confirmed cases / Hospital beds> | <Deaths / Hospital beds>)

*note*: It always shows Portugal, Spain, Italy, South Korea, Iran and Germany.

```{r hospital.beds, echo=FALSE}

wb.indicator <- function(indicator, name) {
    renamed.100k <- glue("{name} (per 100,000 people)")
    
    dta.original <- wb(indicator = indicator)
    
    dta.norm <- dta.original %>%
        mutate(value = value * 100, value,
               indicator = renamed.100k)
    
    for (ix.name in names(convert.names)) {
        dta.norm <- dta.norm %>%
            arrange(country) %>%
            ungroup %>%
            mutate(country = replace(country, country == ix.name, convert.names[[ix.name]]))    
    }
    
    dta <- dta.norm %>% 
        group_by(country) %>%
        arrange(desc(date)) %>%
        summarise(date = first(date),
                  value = first(value))
    
    dta.vs.cases <- left_join(death.vs.cases, dta, by = 'country') %>%
        mutate(ratio.confirmed = cases.confirmed / value) %>%
        mutate(ratio.death = cases.death / value)
    
    return(dta.vs.cases)
}
# death.vs.cases$country[!death.vs.cases$country %in% dta$country]
# dta.norm$country %>% unique %>% sort

beds.vs.cases <- wb.indicator('SH.MED.BEDS.ZS', 'Hospital beds')
```

```{r hospital.beds.func, echo=FALSE}
plot.what.vs.cases <- function(data,
                               country.filter = c(), 
                               always.include = c(),
                               what = 'Hospital beds') {
    my.tbl <- data %>%
        filter(country %in% (c(always.include, country.filter) %>% unique))
    return(
        my.tbl %>%
            ggplot(aes(x = cases.confirmed, y = value, color = country)) +
                    geom_point(aes(size = ratio.confirmed), alpha = .4) +
                    geom_label_repel(aes(label = paste0(country, 
                                                        ' (', percent(ratio.confirmed, accuracy = .1), '/', 
                                                              percent(ratio.death, accuracy = .1), ')'),
                                         fill = country),
                                         na.rm = TRUE,
                                         alpha = .8,
                                         color = 'white',
                                         size = 3,
                                         segment.alpha = .4,
                                         segment.colour = 'black',
                                         force = 2) +
                    expand_limits(x = ceiling(max(my.tbl %>% pull(cases.confirmed))),
                                  y = ceiling(max(my.tbl %>% pull(value)))) +
                    scale_color_viridis(discrete = TRUE, end = .85) +
                    scale_fill_viridis(discrete = TRUE, end = .85) +
                    labs(x = 'Confirmed Cases per 100k population', 
                         y = glue('{what} per 100k population'), 
                         title = glue('{what} vs. Cases per 100k population'), 
                         subtitle = glue('percentage shows the {tolower(what)} per confirmed cases and deaths, respectively'),
                         caption = last.date.string) +
                    theme_minimal() +
                    theme(legend.position = 'none')
        )
}
plot.beds.vs.cases <- function(...) { 
    plot.what.vs.cases(..., 
                       data = beds.vs.cases, 
                       what = 'Hospital beds', 
                       always.include = countries.always.include)
}
```

## EU

```{r eu.beds, echo=FALSE}
plot.beds.vs.cases(countries.eu)
```

## Non-EU Europe

```{r non.eu.europe.beds, echo=FALSE}
plot.beds.vs.cases(countries.europe.non.eu)
```

## Non-Europe

```{r non.europe.beds, echo=FALSE}
plot.beds.vs.cases(countries.non.europe)
```

## CPLP

```{r cplp.beds, echo=FALSE}
plot.beds.vs.cases(countries.cplp)
```

## Data

```{r beds.vs.cases, echo=FALSE}
beds.vs.cases %>% 
    filter(country %in% countries.extended) %>%
    mutate(Country = country,
           Population = format(population, big.mark = ','),
           'Deaths per 100k' = round(cases.death, digits = 2),
           'Cases per 100k' = round(cases.confirmed, digits = 2),
           '(source year)' = as.numeric(date),
           'Beds per 100k' = value,
           'Cases per Beds' = percent(cases.confirmed / value, accuracy = .01),
           'Deaths per Beds' = percent(cases.death / value, accuracy = .01)) %>%
    ungroup %>%
    select(Country,
           Population, 
           'Beds per 100k', 
           '(source year)' , 
           'Cases per 100k', 
           'Deaths per 100k', 
           'Cases per Beds',
           'Deaths per Beds') %>%
    datatable()
```

# Nurses and Midwifes *(per 100k population)*  {.tabset .tabset-fade .tabset-pills}

```{r nurses, echo=FALSE}
nurses.vs.cases <- wb.indicator('SH.MED.NUMW.P3', 'Nurses and midwives')
```

```{r nurses.func, echo=FALSE}
plot.nurses.vs.cases <- function(...) { plot.what.vs.cases(..., data = nurses.vs.cases, what = 'Nurses and Midwifes') }
```

## EU

```{r eu.nurses, echo=FALSE}
plot.nurses.vs.cases(countries.eu)
```

## Non-EU Europe

```{r non.eu.europe.nurses, echo=FALSE}
plot.nurses.vs.cases(countries.europe.non.eu)
```

## Non-Europe

```{r non.europe.nurses, echo=FALSE}
plot.nurses.vs.cases(countries.non.europe)
```

## CPLP

```{r cplp.nurses, echo=FALSE}
plot.nurses.vs.cases(countries.cplp)
```


## Data

```{r nurses.vs.cases, echo=FALSE}
nurses.vs.cases %>% 
    filter(country %in% countries.extended) %>%
    mutate(Country = country,
           Population = format(population, big.mark = ','),
           'Deaths per 100k' = round(cases.death, digits = 2),
           'Cases per 100k' = round(cases.confirmed, digits = 2),
           '(source year)' = as.numeric(date),
           'Nurses per 100k' = value,
           'Cases per Nurses' = percent(cases.confirmed / value, accuracy = .01),
           'Deaths per Nurses' = percent(cases.death / value, accuracy = .01)) %>%
    ungroup %>%
    select(Country,
           Population, 
           'Nurses per 100k', 
           '(source year)' , 
           'Cases per 100k', 
           'Deaths per 100k', 
           'Cases per Nurses',
           'Deaths per Nurses') %>%
    datatable()
```

# Physicians *(per 100k population)*  {.tabset .tabset-fade .tabset-pills}


```{r physicians, echo=FALSE}
physicians.vs.cases <- wb.indicator('SH.MED.PHYS.ZS', 'Physicians')
```

```{r physicians.func, echo=FALSE}
plot.physicians.vs.cases <- function(...) { plot.what.vs.cases(..., data = physicians.vs.cases, what = 'Physicians') }
```

## EU

```{r eu.physicians, echo=FALSE}
plot.physicians.vs.cases(countries.eu)
```

## Non-EU Europe

```{r non.eu.europe.physicians, echo=FALSE}
plot.physicians.vs.cases(countries.europe.non.eu)
```

## Non-Europe

```{r non.europe.physicians, echo=FALSE}
plot.physicians.vs.cases(countries.non.europe)
```

## CPLP

```{r cplp.physicians, echo=FALSE}
plot.physicians.vs.cases(countries.cplp)
```

## Data

```{r physicians.vs.cases, echo=FALSE}
physicians.vs.cases %>% 
    filter(country %in% countries.extended) %>%
    mutate(Country = country,
           Population = format(population, big.mark = ','),
           'Deaths per 100k' = round(cases.death, digits = 2),
           'Cases per 100k' = round(cases.confirmed, digits = 2),
           '(source year)' = as.numeric(date),
           'Physicians per 100k' = value,
           'Cases per Physicians' = percent(cases.confirmed / value, accuracy = .01),
           'Deaths per Physicians' = percent(cases.death / value, accuracy = .01)) %>%
    ungroup %>%
    select(Country,
           Population, 
           'Physicians per 100k', 
           '(source year)' , 
           'Cases per 100k', 
           'Deaths per 100k', 
           'Cases per Physicians',
           'Deaths per Physicians') %>%
    datatable()
```


\- - - - - - - - The end. - - - - - - - -

```{r handle.data, eval=FALSE, include=FALSE}
anim.confirmed.vs.death <- last.week %>%
    filter(country %in% c('Italy', 'Portugal')) %>%
    group_by(country, date) %>%
    mutate(cases.death = cumulative.cases.death / population * 100000,
           cases.confirmed = cumulative.cases.confirmed / population * 100000,
           ratio = if_else(cumulative.cases.confirmed == 0, 0, cumulative.cases.death / cumulative.cases.confirmed)) %>%
    select(country, date, cases.confirmed, cases.death, cumul.confirmed = cumulative.cases.confirmed, cumul.death = cumulative.cases.death, ratio, population) %>%
    arrange(desc(date))

anim.confirmed.vs.death
```

```{r animate, eval=FALSE, include=FALSE}
anim <- anim.confirmed.vs.death %>%
    arrange(date) %>%
    #filter(date == '2020-03-24') %>%
    ggplot(aes(x = cases.confirmed, y = cases.death, color = country)) +
        geom_point(aes(size = population), alpha = .4) +
        geom_label_repel(aes(label = paste0(country, ' (', percent(ratio, accuracy = .1), ')'),
                     fill = country),
                     na.rm = TRUE,
                     alpha = .8,
                     color = 'white',
                     size = 3,
                     segment.alpha = .4,
                     segment.colour = 'black',
                     force = 2) +
        expand_limits(x = ceiling(max(anim.confirmed.vs.death %>% pull(cases.confirmed))),
                          y = ceiling(max(anim.confirmed.vs.death %>% pull(cases.death)))) +
            labs(x = 'Confirmed Cases per 100k population', 
                 y = 'Deaths per 100k population', 
                 title = 'Deaths vs. Cases per 100k population -- {format(frame_along, \'%B\ %d\')}', 
                 subtitle = 'percentage is the death rate per confirmed cases and size represents size of country by population',
                 caption = '??') +
            theme_minimal() +
            theme(legend.position = 'none') +
    transition_time(date) +
    enter_drift()
```

```{r show_anim, message=FALSE, eval=FALSE, include=FALSE}
animate(anim, end_pause = 30)
```
