---
title: "World covid-19 Analysis"
output: 
    html_document:
        toc: true
        dev: 'svg'
params:
    death.start: !r 10
    confirmed.start: !r 100
    duplicate.every: !r 2
    last.days: !r 12
---

```{r, child='_header.md'}
```

```{r, include=FALSE, eval=FALSE}
rmarkdown::render('index.Rmd', quiet = TRUE)
rmarkdown::render('index.Rmd')
```

```{r, include=FALSE, eval=FALSE}
rmarkdown::clean_site()
rmarkdown::render_site(quiet = TRUE)
#rmarkdown::render_site()
```

```{r, eval=FALSE, include=FALSE}
BiocManager::install('averissimo/covid19.pt.data')
BiocManager::install('averissimo/covid19.de.data')
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, collapse = TRUE, fig.width = 10, fig.height = 8)
library(dplyr)
library(purrr)
library(tibble)
library(plotly)
library(viridis) # color pallete
library(ggrepel)
library(zoo)
library(glue)
library(DT)
library(loose.rock)
library(anytime)
library(scales)
library(futile.logger)
library(wbstats) # world bank data
library(reshape2)
library(eurostat)

devtools::load_all()

Sys.setlocale('LC_TIME', 'en_GB.UTF-8')
.Last.value <- flog.layout(layout.format('~m'))

theme_set(
  theme_minimal() +
    theme(legend.position = "none")
  )
```

```{r, include=FALSE}
BiocManager::install('averissimo/covid19.de.data', ask = FALSE, update = TRUE)
BiocManager::install('averissimo/covid19.pt.data', ask = FALSE, update = TRUE)
devtools::load_all()
```


```{r, message=FALSE, warning=FALSE, include=FALSE}
eucdc.dat  <- download.eucdc.data()
jh.dat     <- download.john.hopkins()

if (max(jh.dat$data$date) > max(eucdc.dat$data$date)) {
  world.dat <- jh.dat  
  
#  state.ix <- eucdc.dat$data$state %>% unique()
#  state.ix[!state.ix %in% (jh.dat$data$state %>% unique())]  
} else {
  world.dat <- eucdc.dat
  
#  state.ix <- jh.dat$data$state %>% unique()
#  state.ix[!state.ix %in% (eucdc.dat$data$state %>% unique())]  
}

it.dat.raw <- download.it.data(by.state = FALSE)
pt.dat.raw <- download.pt.data()
de.dat.raw <- download.de.data()
es.dat.raw <- download.es.data()

latest <- function(target, dat, one.dat) { 
  flog.info('%s', target)

  internal.fun <- function(my.dat) {
    output <- my.dat$data %>% filter(state == target) %>% group_by(state, type) %>% top_n(1, date) 
    
    output.cases <- output %>% filter(type == 'confirmed')
    output.deaths <- output %>% filter(type == 'death')  
  
    flog.info('  %s', my.dat$source)
    flog.info('     cases: %d (%s)', output.cases %>% pull(cases), output.cases %>% pull(date))
    flog.info('    deaths: %d (%s)', output.deaths %>% pull(cases), output.deaths %>% pull(date))
  }
  
  internal.fun(dat)
  internal.fun(one.dat)
}

latest('Spain', world.dat, es.dat.raw)
latest('Portugal', world.dat, pt.dat.raw)
latest('Germany', world.dat, de.dat.raw)
latest('Italy', world.dat, it.dat.raw)


world.dat.norm <- world.dat
world.dat.norm <- replace.country.data(world.dat.norm, es.dat.raw, 'Spain')
world.dat.norm <- replace.country.data(world.dat.norm, pt.dat.raw, 'Portugal')
world.dat.norm <- replace.country.data(world.dat.norm, it.dat.raw, 'Italy')
world.dat.norm <- replace.country.data(world.dat.norm, de.dat.raw, 'Germany')

# Keep only confirmed and death numbers
world.dat.norm$data <- world.dat.norm$data %>% filter(type %in% c('confirmed', 'death'))

source.by       <- world.dat.norm$source

flog.info('')
flog.info('Post normalization')

latest('Portugal', world.dat.norm, pt.dat.raw)
latest('Germany', world.dat.norm, de.dat.raw)
latest('Italy', world.dat.norm, it.dat.raw)

last.date        <- source.by
last.date.string <- paste0('Latest date from ', last.date)

my.corona.original <- world.dat.norm$data %>% 
    filter(state != 'Diamond Princess') %>%
    ungroup() %>%
    mutate(state.code = tolower(state.code)) %>% 
    arrange(date, state, type) %>%
    group_by(state, type)

my.corona <- my.corona.original

my.corona <- mutate(my.corona, 
                    state.data = factor(state, 
                                        levels = my.corona %>% filter(type == 'confirmed') %>% summarise(cases = max(cumul)) %>% arrange(cases) %>% pull(state)))
```

```{r}
#last.date          <- format(max(my.corona.original$date), '%Y/%m/%d')
#last.date.string   <- paste0('Latest date from ', last.date, ' (source: ', source.by, ')')
region.code        <- 'Country'
region.code.plural <- 'Countries'
region.all         <- 'World'

countries          <- c('Italy', 'Spain','Germany', 'Portugal', 'USA', 'France', 
                        'Switzerland', 'Czech Republic', 'Sweden', 'United Kingdom') %>% sort

countries.eu <- c('Italy', 'Spain','Germany', 'Portugal', 'France',  
                  'United Kingdom', 'Poland', 'Greece', 'Croatia', 'Slovenia', 
                  'Hungary', 'Slovakia', 'Ireland', 'Luxembourg', 'Finland', 
                  'Sweden', 'Denmark', 'Romania', 'Netherlands', 'Belgium', 
                  'Czech Republic', 'Austria', 'Bulgaria', 'Lithuania', 'Latvia', 
                  'Malta', 'Estonia', 'Cyprus')
countries.cplp <- c('Angola', 'Mozambique', 'Cape Verde', 'Timor Leste', 
                    'Brazil'
                    , 'Macao SAR, China', 'Guinea Bissau', 'Sao Tome and Principe' # not in corona data
                    )
countries.europe.non.eu <- c('Norway', 'Switzerland', 'Ukraine', 'Serbia', 'Russia', 
                             'Albania', 'Iceland', 'Turkey', 'Belarus')
countries.non.europe <- c('USA', 'China', 'Canada', 
                          'Australia', 'New Zealand', 
                          'South Korea', 'Japan', 'Argentina', 'Chile', 'Thailand',
                          'Iran', 'China', countries.cplp)

countries.always.include <- c('Portugal', 'USA', 'Germany', 'Spain', 'Italy', 'South Korea')
countries.extended <- c(countries.eu, countries.europe.non.eu, countries.cplp, countries.non.europe, countries.always.include) %>% unique
```

```{r}
after.100.dat = list(all = filter.after(my.corona, params$death.start, params$confirmed.start)) # my.corona.after.100
after.100.dat$confirmed = ommit.data(after.100.dat$all, 'confirmed', countries) # my.corona.after.100.confirmed.dat
after.100.dat$death = ommit.data(after.100.dat$all, 'death', countries) # my.corona.after.100.death.dat

last.days.dat <- filter.last.days(my.corona, params$last.days) # my.corona.last.days
last.week.dat <- filter.last.days.cumulative(my.corona, 4) # last.week
death.vs.cases <- filter.death.vs.cases(last.week.dat) %>% filter(state %in% countries.extended) # death.vs.cases

my.world = list()
my.world$main <- my.corona %>% 
    group_by(type, date) %>%
    summarise(cases = sum(cases), population = sum(population)) %>%
    arrange(date) %>%
    group_by(type) %>% 
    mutate(state = region.all,
           state.code = region.all,
           cumul = cumsum(cases),
           population = max(population, na.rm = TRUE))

# Only for world population
my.world$main$population <- run.cache(wb, indicator = "SP.POP.TOTL", show.message = FALSE) %>% 
  filter(country == 'World') %>% 
  top_n(1, date) %>%
  pull(value) %>% 
  pluck(1)

my.world$last.days <- filter.last.days(my.world$main, params$last.days)
my.world$last.week <- filter.last.days.cumulative(my.world$main, 4)
```

```{r, eval=FALSE, include=FALSE}
my.corona %>% 
  top_n(3, date) %>% 
  dcast(state + date ~ type, value.var = 'cases')

my.corona %>% 
  filter(cases != 0) %>% 
  top_n(2, date) %>% 
  dcast(state + date ~ type, value.var = 'cases')
```


```{r, child='_data.Rmd'}
```

```{r, child='_main_plots.Rmd'}
```


```{r, child='_world.Rmd'}
```


```{r, child='_main.Rmd'}
```






























## EU

```{r}
death.vs.cases.plot(death.vs.cases, state.filter = countries.eu, always.include = countries.always.include)
```

## Non-EU Europe

```{r}
death.vs.cases.plot(death.vs.cases, state.filter = countries.europe.non.eu, always.include = countries.always.include)
```

## Non-Europe

```{r}
death.vs.cases.plot(death.vs.cases, state.filter = countries.non.europe, always.include = countries.always.include)
```

## CPLP

```{r}
death.vs.cases.plot(death.vs.cases, state.filter = countries.cplp, always.include = countries.always.include)
```

## Data

```{r}
death.vs.cases.data(death.vs.cases) %>% datatable(rownames = FALSE)
```

## Number of Confirmed Cases

```{r}
my.corona.original %>% filter(state %in% countries.extended) %>% cases.plot('confirmed')
```

## Number of Deaths

```{r}
my.corona.original %>% filter(state %in% countries.extended) %>% cases.plot('death')
```

# Hospital beds vs. Cases *(per 100k population)* {.tabset .tabset-fade .tabset-pills}

This visualization shows the number of cases/deaths per 100k of the population.

* **x axis**: Confirmed cases per 100k population
* **y axis**: Hospital beds per 100k population
* **size of data point**: Ratio of Confirmed cases / Hospital beds
* **label**: `r region(1)` (<Confirmed cases / Hospital beds> | <Deaths / Hospital beds>)

*note*: It always shows Portugal, Spain, Italy, South Korea, Iran and Germany.

```{r}
beds.vs.cases <- wb.indicator(death.vs.cases, 'SH.MED.BEDS.ZS', 'Hospital beds', eu.convert.names)
```

```{r}
plot.beds.vs.cases <- function(...) { 
    plot.what.vs.cases(..., 
                       data = beds.vs.cases, 
                       what = 'Hospital beds', 
                       always.include = countries.always.include)
}
```

## EU

```{r}
plot.beds.vs.cases(countries.eu)
```

## Non-EU Europe

```{r}
plot.beds.vs.cases(countries.europe.non.eu)
```

## Non-Europe

```{r}
plot.beds.vs.cases(countries.non.europe)
```

## CPLP

```{r}
plot.beds.vs.cases(countries.cplp)
```

## Data

```{r}
something.vs.cases.data(beds.vs.cases, 'Beds') %>% datatable(rownames = FALSE)
```

# Nurses and Midwifes *(per 100k population)*  {.tabset .tabset-fade .tabset-pills}

```{r}
nurses.vs.cases <- wb.indicator(death.vs.cases, 'SH.MED.NUMW.P3', 'Nurses and midwives', eu.convert.names)
```

```{r}
plot.nurses.vs.cases <- function(...) { plot.what.vs.cases(..., data = nurses.vs.cases, what = 'Nurses and Midwifes') }
```

## EU

```{r}
plot.nurses.vs.cases(countries.eu)
```

## Non-EU Europe

```{r}
plot.nurses.vs.cases(countries.europe.non.eu)
```

## Non-Europe

```{r}
plot.nurses.vs.cases(countries.non.europe)
```

## CPLP

```{r}
plot.nurses.vs.cases(countries.cplp)
```


## Data

```{r}
something.vs.cases.data(nurses.vs.cases, 'Nurses') %>% datatable(rownames = FALSE)
```

# Physicians *(per 100k population)*  {.tabset .tabset-fade .tabset-pills}


```{r}
physicians.vs.cases <- wb.indicator(death.vs.cases, 'SH.MED.PHYS.ZS', 'Physicians', eu.convert.names)
```

```{r}
plot.physicians.vs.cases <- function(...) { plot.what.vs.cases(..., data = physicians.vs.cases, what = 'Physicians') }
```

## EU

```{r}
plot.physicians.vs.cases(countries.eu)
```

## Non-EU Europe

```{r}
plot.physicians.vs.cases(countries.europe.non.eu)
```

## Non-Europe

```{r}
plot.physicians.vs.cases(countries.non.europe)
```

## CPLP

```{r}
plot.physicians.vs.cases(countries.cplp)
```

## Data

```{r}
something.vs.cases.data(physicians.vs.cases, 'Nurses') %>% datatable(rownames = FALSE)
```


\- - - - - - - - The end. - - - - - - - -

```{r, eval=FALSE, include=FALSE}
anim.confirmed.vs.death <- last.week %>%
    filter(country %in% c('Italy', 'Portugal')) %>%
    group_by(country, date) %>%
    mutate(cases.death = cumulative.cases.death / population * 100000,
           cases.confirmed = cumulative.cases.confirmed / population * 100000,
           ratio = if_else(cumulative.cases.confirmed == 0, 0, cumulative.cases.death / cumulative.cases.confirmed)) %>%
    select(country, date, cases.confirmed, cases.death, cumul.confirmed = cumulative.cases.confirmed, cumul.death = cumulative.cases.death, ratio, population) %>%
    arrange(desc(date))

anim.confirmed.vs.death
```

```{r, eval=FALSE, include=FALSE}
anim <- anim.confirmed.vs.death %>%
    arrange(date) %>%
    #filter(date == '2020-03-24') %>%
    ggplot(aes(x = cases.confirmed, y = cases.death, color = country)) +
        geom_point(aes(size = population), alpha = .4) +
        geom_label_repel(aes(label = paste0(country, ' (', percent(ratio, accuracy = .1), ')'),
                     fill = country),
                     na.rm = TRUE,
                     alpha = .8,
                     color = 'white',
                     size = 3,
                     segment.alpha = .4,
                     segment.colour = 'black',
                     force = 2) +
        expand_limits(x = ceiling(max(anim.confirmed.vs.death %>% pull(cases.confirmed))),
                          y = ceiling(max(anim.confirmed.vs.death %>% pull(cases.death)))) +
            labs(x = 'Confirmed Cases per 100k population', 
                 y = 'Deaths per 100k population', 
                 title = 'Deaths vs. Cases per 100k population -- {format(frame_along, \'%B\ %d\')}', 
                 subtitle = 'percentage is the death rate per confirmed cases and size represents size of {region(1, TRUE)} by population' %>% glue,
                 caption = '??') +
            theme_minimal() +
            theme(legend.position = 'none') +
    transition_time(date) +
    enter_drift()
```

```{r, message=FALSE, eval=FALSE, include=FALSE}
animate(anim, end_pause = 30)
```
