---
title: "Portugal Mortality data"
output: 
  html_document:
    dev: 'svg'
    toc: true
---

```{r, eval=FALSE, include=FALSE}
rmarkdown::render('mortality.Rmd', output_file = 'mortality.md')
rmarkdown::render('mortality.Rmd', output_file = 'mortality.html')
```

```{r, eval=FALSE, include=FALSE}
rmarkdown::clean_site()
rmarkdown::render_site(quiet = TRUE)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  fig.height = 10,
  fig.width = 10
)
devtools::load_all()
library(pdftools)
library(rvest)
library(dplyr)
library(tibble)
library(DT)
library(anytime)
library(glue)
library(reshape2)
library(zoo)
library(xml2)
library(ggplot2)
library(ggrepel)
library(plotly)

Sys.setlocale('LC_TIME', 'en_GB.UTF-8')
```

Updated on    `r format(Sys.time(), '%A, %d %B %Y (%d %Y-%m-%d %H:%m %Z)')`

## Description & Code {.tabset .tabset-fade .tabset-pills}

This analysis retrieves data directly from the Portuguese Health ministry death certificate site ([SICO](https://evm.min-saude.pt/)) and makes an yearly comparison.

### Hide source code

### Show source code

#### Day analysis for every year

```{r, echo=TRUE}
# Download from DGS
json <- download.evm() %>% rjson::fromJSON()

# Process into multiple columns
dat <- sapply(1 + seq_along(json$x$data[-1]), function(ix) {
  my.year <- json$x$data[[ix]]
  my.year[sapply(my.year, is.null)] <- NA
  tmp <- tibble::tibble(!!as.name(2009 + ix - 2) := unlist(my.year, recursive=FALSE))
  # current day info may not be finalized
  if (2009 + ix - 2 == format(Sys.Date(), "%Y")) {
    tmp[which(is.na(tmp[,1]) %>% as.vector)[1] - 1, 1] <- NA
  }
  return(tmp)
})

# Merge data
mortality <- dplyr::bind_cols(day = json$x$data[[1]], dat) %>% 
  mutate(day = factor(day, levels = day))

# Add some statistics (mean, sd, max, min)
mortality.with.stats <- mortality %>% 
  replace(is.na(.), NA) %>%
  dplyr::rowwise() %>% 
  mutate(meanPre2020 = mean(c_across(cols = -c("2020", 'day')), na.rm = TRUE),
         mean = mean(c_across(cols = -c("meanPre2020", 'day')), na.rm = TRUE),
         sdPre2020 = sd(c_across(cols = -c("2020","meanPre2020", "mean", 'day')), na.rm = TRUE),
         sd = sd(c_across(cols = -c("meanPre2020", "mean", "sdPre2020", 'day')), na.rm = TRUE)) %>% 
  mutate(maxPre2020 = max(c_across(cols = colnames(.)[colnames(mortality) != 2020][grepl('^[0-9]+$', colnames(.))]), na.rm = TRUE),
         max = max(c_across(cols = colnames(.)[grepl('^[0-9]+$', colnames(.))]), na.rm = TRUE),
         minPre2020 = min(c_across(cols = colnames(.)[colnames(mortality) != 2020][grepl('^[0-9]+$', colnames(.))]), na.rm = TRUE),
         min = min(c_across(cols = colnames(.)[grepl('^[0-9]+$', colnames(.))]), na.rm = TRUE)) %>% 
  dplyr::ungroup()
```

#### Poisson baseline distribution vs. 2020

```{r, echo=TRUE}
pois.levels <- c('2021-lambda.5',
                 '2020-lambda.5', 
                 'pre2020-lambda', 
                 'pre2020-lambda.5', 
                 'pre2020-ic95.5', 
                 'pre2020-ic99.5')
pois.labels <- c('2020 (smooth regression)',  # this is new
                 '2021 (smooth regression)',  # this is new
                 #
                 '2020',
                 '2021',
                 'Confidence interval 95% 2015-19',
                 'Confidence interval 99% 2015-19',
                 'Base line for 2015-19 (non-averaged)',
                 'Base line for 2015-19')

average.days <- 7

mortality.pois <- mortality %>% 
  reshape2::melt(id.vars = c('day'), variable.name = 'year') %>% 
  filter(!is.na(value)) %>% 
  mutate(year = as.numeric(as.character(year))) %>% 
  #filter(year >= 2015) %>% 
  mutate(group = if_else(year >= 2020, as.character(year), 'pre2020')) %>% 
  group_by(group, day) %>% 
  dplyr::arrange(day, group)

mortality.pois.transf <- mortality.pois  %>% 
  summarise(lambda = mean(value, na.rm = TRUE), .groups = 'keep') %>% 
  mutate(ic95 = qpois(.95, lambda = lambda, lower.tail = TRUE),
         ic99 = qpois(.99, lambda = lambda, lower.tail = TRUE)) %>% 
  dplyr::group_by(group) %>% 
  mutate(lambda.5 = zoo::rollapply(lambda, width = average.days, FUN = function(x) {mean(x,na.rm=TRUE)}, partial = TRUE, fill = NA, align = 'center'),
         ic95.5 = qpois(.95, lambda = lambda.5, lower.tail = TRUE),
         ic99.5 = qpois(.99, lambda = lambda.5, lower.tail = TRUE)) %>% 
  #
  reshape2::melt(id.vars = c('group', 'day')) %>% 
  mutate(new.group = paste0(group, '-', variable)) %>% 
  filter(new.group %in% pois.levels) %>% 
  mutate(ymin = min(value),
         new.group = factor(new.group, levels = pois.levels)) %>% 
  arrange(day, group)
  
mortality.plot.poisson <- mortality.pois.transf %>% 
  ggplot(aes(x = day, group = new.group, color = new.group)) +
  geom_ribbon(aes(y = value, fill= new.group, ymin = ymin - abs(ymin * .05), ymax = value), alpha = .4, position = 'identity', 
              data = mortality.pois.transf %>% filter(new.group %in% c('pre2020-lambda.5','pre2020-ic99.5','pre2020-ic95.5')),
              show.legend = FALSE) +
  geom_line(aes(y = value, color = new.group)) +
  geom_line(aes(y = value, color = new.group), data = mortality.pois.transf %>% filter(group == '2021'), size = 1.5) +
  geom_line(aes(y = value, color = new.group), data = mortality.pois.transf %>% filter(group == '2020'), size = 1.5) +
  geom_smooth(aes(y = value, color = '2021 (smooth regression)'), alpha = .4, data = mortality.pois.transf %>% filter(group == '2021'), span = 0.1, method = 'loess', formula = y ~ x, na.rm = TRUE, show.legend = FALSE, se = FALSE) +
  geom_smooth(aes(y = value, color = '2020 (smooth regression)'), alpha = .4, data = mortality.pois.transf %>% filter(group == '2020'), span = 0.1, method = 'loess', formula = y ~ x, na.rm = TRUE, show.legend = FALSE, se = FALSE) +
  #geom_line(aes(y = lambda.5, color = new.grou~p), data = mortality.pois.transf %>% filter(group == 'pre2020')) +
  #geom_line(aes(y = ic95.5), color = 'purple', data = mortality.pois.transf %>% filter(group == 'pre2020')) +
  #geom_line(aes(y = ic99.5), color = 'orange', data = mortality.pois.transf %>% filter(group == 'pre2020')) +
  #geom_line(aes(y = lambda.5), data = mortality.pois.transf %>% filter(group == '2020')) +
  scale_x_discrete(breaks = function(ix) {
      ix[c(FALSE, TRUE, TRUE, TRUE, TRUE)] <- FALSE
      return(ix)
    }) +
  scale_color_discrete("", labels = pois.labels) +
  labs(title = '2020 and 2021 vs Baseline using Poisson distribution', 
       subtitle = 'Data calculated with centered average of {average.days} days' %>% glue::glue(),
       caption = "Data retrieved on {format(Sys.Date(), '%B %d, %Y')}" %>% glue::glue()) +
  xlab('Day of year') +
  ylab('Number of deaths') +
  theme_minimal() +
  theme(legend.key = element_rect(fill = "white", colour = "white"), legend.position = 'top', axis.text.x = element_text(angle = 45, hjust = 1, size = 5))
```

Sample plot

```{r sample_plot, eval=FALSE, echo=TRUE}
show.year.mortality(mortality.with.stats, seq(2019, 2020))
```

#### Yearly stats from 2009-2020

```{r, echo=TRUE}
yearly.stats <- mortality[c(rep(0, which(mortality$day == 'Mar-15') - 1), seq(from = which(mortality$day == 'Mar-15'), nrow(mortality)) > 0) & as.vector(!is.na(mortality[,'2020'])),] %>% 
  rowwise() %>% 
  replace(is.na(.), NA) %>%
  #mutate(maxPre2020    = max(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE),
  #       minPre2020    = min(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE),
  #       meanPre2020   = mean(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE),
  #       medianPre2020 = median(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE)) %>% 
  reshape2::melt(id.vars = 'day') %>% 
  dplyr::group_by(variable) %>% 
  dplyr::summarise(mean = mean(value, na.rm = TRUE), 
            sd = sd(value, na.rm = TRUE),
            sum = sum(value, na.rm = TRUE),
            median = median(value, na.rm = TRUE),
            max = max(value, na.rm = TRUE),
            min = min(value, na.rm = TRUE),
            .groups = 'rowwise')
```

```{r, echo=TRUE}
yearly.comp <- sapply(seq(nrow(yearly.stats)), function(ix) { yearly.stats$sum - yearly.stats$sum[ix]}) %>%
  data.frame()
yearly.comp[upper.tri(yearly.comp)] <- NA
colnames(yearly.comp) <- yearly.stats$variable
yearly.comp <- bind_cols(year = yearly.stats$variable, yearly.comp)
```

```{r, echo=TRUE}
yearly.comp.rel <- sapply(seq(nrow(yearly.stats)), function(ix) { yearly.stats$sum / yearly.stats$sum[ix]}) %>%
  data.frame()
yearly.comp.rel[upper.tri(yearly.comp.rel)] <- NA
colnames(yearly.comp.rel) <- yearly.stats$variable
yearly.comp.rel <- bind_cols(year = yearly.stats$variable, yearly.comp.rel)
```

#### Yearly stats calculated _(max, min, mean, median and mean +/- StDev)_

```{r, echo=TRUE}
yearly.stats.2 <- mortality[c(rep(0, which(mortality$day == 'Mar-15') - 1), seq(from = which(mortality$day == 'Mar-15'), nrow(mortality)) > 0) & as.vector(!is.na(mortality[,'2020'])),] %>% 
  rowwise() %>% 
  replace(is.na(.), NA) %>%
  mutate(maxPre2020    = max(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE),
         minPre2020    = min(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE),
         meanPre2020   = mean(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE),
         meanPlusStDevPre2020   = mean(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE) + sd(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE),
         meanMinusStDevPre2020   = mean(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE) - sd(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE),
         medianPre2020 = median(c_across(colnames(.)[!grepl('(^2020)',colnames(.)) & grepl('^[0-9]+$',colnames(.))]), na.rm = TRUE)) %>% 
  reshape2::melt(id.vars = 'day') %>% 
  dplyr::group_by(variable) %>% 
  dplyr::summarise(mean = mean(value, na.rm = TRUE), 
            sd = sd(value, na.rm = TRUE),
            sum = sum(value, na.rm = TRUE),
            median = median(value, na.rm = TRUE),
            max = max(value, na.rm = TRUE),
            min = min(value, na.rm = TRUE),
            .groups = 'rowwise') %>% 
  filter(grepl('202[0-9]$', variable)) %>% 
  mutate(variable = factor(variable, levels = c( 'medianPre2020', 'minPre2020', 'meanMinusStDevPre2020', 'meanPre2020', 'meanPlusStDevPre2020', 'maxPre2020', '2020'))) %>% 
  arrange(variable)
```

```{r, echo=TRUE}
yearly.comp.rel.2 <- sapply(seq(nrow(yearly.stats.2)), function(ix) { yearly.stats.2$sum / yearly.stats.2$sum[ix]}) %>%
  data.frame()
yearly.comp.rel.2[upper.tri(yearly.comp.rel.2)] <- NA
colnames(yearly.comp.rel.2) <- yearly.stats.2$variable
yearly.comp.rel.2 <- bind_cols(year = yearly.stats.2$variable, yearly.comp.rel.2)
```

#### External causes data

```{r, echo=TRUE}
external.raw <- download.causas.externas()

json.ext <- external.raw$json %>% 
  purrr::map(function(ix) {rjson::fromJSON(ix)})

external.causes <- lapply(seq_along(external.raw$year), function(ix.year) {
  # Process into multiple columns
  json.tmp <- json.ext[[ix.year]]
  dat <- sapply(1 + seq_along(json.tmp$x$data[-1]), function(ix) {
    my.colnames <- xml2::read_xml(json.tmp$x$container) %>% 
      rvest::html_nodes('th') %>%
      gsub(' ?</?th> ?', '', .)
    
    my.year <- json.tmp$x$data[[ix]]
    my.year[sapply(my.year, is.null)] <- NA
    tmp <- tibble::tibble(!!as.name(my.colnames[ix]) := unlist(my.year, recursive=FALSE))
    return(tmp)
  })
  dat$year <- rep(external.raw$year[ix.year], length(dat[[1]]))
  
  # Merge data
  return(dplyr::bind_cols(day = json.tmp$x$data[[1]], dat) %>% 
    mutate(day = factor(day, levels = json.ext[[1]]$x$data[[1]])))
}) %>% 
  bind_rows %>% 
  dplyr::relocate(year, .before = day) %>% 
  mutate(year = factor(year))
```

#### External causes plots

```{r, echo=TRUE}
external.causes.plot.covid <- external.causes %>% 
  filter(day %in% levels(external.causes$day)[levels(external.causes$day) %in% (external.causes %>% filter(year == '2020') %>% pull(day))]) %>% 
  filter(day %in% levels(external.causes$day)[seq(from = which(levels(external.causes$day) == 'Mar-15'), to = length(levels(external.causes$day)))]) %>% 
  group_by(year) %>% 
  summarise(across(where(is.numeric), ~ sum(.x , na.rm = TRUE)), .groups = 'keep') %>% 
  reshape2::melt(id = 'year') %>% 
  ggplot(aes(y = year, x = variable)) +
  geom_tile(aes(fill = value)) +
  geom_text(aes(label = value), color = 'white') +
  scale_fill_gradient(low = "orange", high = "red", na.value = '#00000020') +
  scale_x_discrete(position = "top") +
  labs(title = "External causes of death from \'March 15 to {format(Sys.Date(), '%B %d')}'" %>% glue::glue(), subtitle = 'All years show the same period', caption = "Accidents, suicides and others. It excludes natural causes or diseases\nData retrieved on {format(Sys.Date(), '%B %d, %Y')}" %>% glue::glue()) + 
  xlab('External cause of death') +
  ylab('Year') +
  theme_minimal() +
  theme(legend.position = 'none')

external.causes.plot.full.year <- external.causes %>% 
  group_by(year) %>% 
  summarise(across(where(is.numeric), ~ sum(.x , na.rm = TRUE)), .groups = 'keep') %>% 
  reshape2::melt(id = 'year') %>% 
  ggplot(aes(y = year, x = variable)) +
  geom_tile(aes(fill = value)) +
  geom_text(aes(label = value), color = 'white') +
  scale_fill_gradient(low = "orange", high = "red", na.value = '#00000020') +
  scale_x_discrete(position = "top") +
  labs(title = 'External causes of death', subtitle = 'All years show the same period', caption = "Accidents, suicides and others. It excludes natural causes or diseases\nData retrieved on {format(Sys.Date(), '%B %d, %Y')}" %>% glue::glue()) + 
  xlab('External cause of death') +
  ylab('Year') +
  theme_minimal() +
  theme(legend.position = 'none')
```

#### Age group data

```{r, echo=TRUE}
age.group.raw <- download.age.groups()

json.age.group <- age.group.raw$json %>% 
  purrr::map(function(ix) {rjson::fromJSON(ix)})

age.group <- lapply(seq_along(age.group.raw$year), function(ix.year) {
  # Process into multiple columns
  json.tmp <- json.age.group[[ix.year]]
  dat <- sapply(1 + seq_along(json.tmp$x$data[-1]), function(ix) {
    my.colnames <- xml2::read_xml(json.tmp$x$container) %>% 
      rvest::html_nodes('th') %>%
      gsub(' ?</?th> ?', '', .)
    
    my.year <- json.tmp$x$data[[ix]]
    my.year[sapply(my.year, is.null)] <- NA
    tmp <- tibble::tibble(!!as.name(my.colnames[ix]) := unlist(my.year, recursive=FALSE))
    return(tmp)
  })
  dat$year <- rep(age.group.raw$year[ix.year], length(dat[[1]]))
  
  # Merge data
  return(dplyr::bind_cols(day = json.tmp$x$data[[1]], dat) %>% 
    mutate(day = factor(day, levels = json.age.group[[1]]$x$data[[1]])))
}) %>% 
  bind_rows %>% 
  dplyr::relocate(year, .before = day) %>% 
  mutate(year = factor(year))
```

#### Plots for age group

```{r, echo=TRUE}
age.group.plot.covid <- age.group %>% 
  filter(day %in% levels(age.group$day)[levels(age.group$day) %in% (age.group %>% filter(year == '2020') %>% pull(day))]) %>% 
  filter(day %in% levels(age.group$day)[seq(from = which(levels(age.group$day) == 'Mar-15'), to = length(levels(age.group$day)))]) %>% 
  rowwise() %>% 
  mutate(total = sum(c_across(-c('year', 'day')))) %>% 
  ungroup() %>% 
  group_by(year) %>% 
  summarise(across(where(is.numeric), ~ sum(.x , na.rm = TRUE)), .groups = 'keep') %>% 
  reshape2::melt(id = c('year', 'total')) %>% 
  mutate(label = factor(paste0(year, '\n(', format(total, big.mark = ' '),')'))) %>% 
  ggplot(aes(y = label, x = variable)) +
  geom_tile(aes(fill = value)) +
  geom_text(aes(label = value), color = 'white') +
  scale_fill_gradient(low = "orange", high = "red", na.value = '#00000020') +
  scale_x_discrete(position = "top") +
  labs(title = "Mortality by age group from \'March 15 to {format(Sys.Date(), '%B %d')}'" %>% glue::glue(), subtitle = 'All years show the same period', caption = "Data retrieved on {format(Sys.Date(), '%B %d, %Y')}" %>% glue::glue()) + 
  xlab('Age group') +
  ylab('Year') +
  theme_minimal() +
  theme(legend.position = 'none')

age.group.plot.full.year <- age.group %>% 
  rowwise() %>% 
  mutate(total = sum(c_across(-c('year', 'day')))) %>% 
  ungroup() %>% 
  group_by(year) %>% 
  summarise(across(where(is.numeric), ~ sum(.x , na.rm = TRUE)), .groups = 'keep') %>% 
  reshape2::melt(id = c('year', 'total')) %>% 
  mutate(label = factor(paste0(year, '\n(', format(total, big.mark = ' '),')'))) %>% 
  ggplot(aes(y = label, x = variable)) +
  geom_tile(aes(fill = value)) +
  geom_text(aes(label = value), color = 'white') +
  scale_fill_gradient(low = "orange", high = "red", na.value = '#00000020') +
  scale_x_discrete(position = "top") +
  labs(title = 'Mortality by age group', subtitle = 'All years show the same period (last year may be incomplete)', caption = "Data retrieved on {format(Sys.Date(), '%B %d, %Y')}" %>% glue::glue()) +  
  xlab('Age group') +
  ylab('Year') +
  theme_minimal() +
  theme(legend.position = 'none')
```

## 2020 vs. Baseline

```{r y2020.alone, warning=FALSE, fig.asp=0.6}
mortality.plot.poisson
```


## Year vs. 2020 {.tabset .tabset-fade .tabset-pills}

Yearly comparison against 2020

### 2019 vs. 2020

```{r y2019}
show.year.mortality(mortality.with.stats, seq(2019, 2020))
```

### 2018 vs. 2020

```{r y2018}
show.year.mortality(mortality.with.stats, c(2018, 2020))
```

### 2017 vs. 2020

```{r y2017}
show.year.mortality(mortality.with.stats, c(2017, 2020))
```

### 2016 vs. 2020

```{r y2016}
show.year.mortality(mortality.with.stats, c(2016, 2020))
```

### 2015 vs. 2020

```{r y2015}
show.year.mortality(mortality.with.stats, c(2015, 2020))
```

### 2014 vs. 2020

```{r y2014}
show.year.mortality(mortality.with.stats, c(2014, 2020))
```

### 2013 vs. 2020

```{r y2013}
show.year.mortality(mortality.with.stats, c(2013, 2020))
```

### 2012 vs. 2020

```{r y2012}
show.year.mortality(mortality.with.stats, c(2012, 2020))
```

### 2011 vs. 2020

```{r y2011}
show.year.mortality(mortality.with.stats, c(2011, 2020))
```

### 2010 vs. 2020

```{r y2010}
show.year.mortality(mortality.with.stats, c(2010, 2020))
```

### 2009 vs. 2020

```{r y2009}
show.year.mortality(mortality.with.stats, c(2009, 2020))
```


## _'March-15 -> `r format(Sys.Date(), '%B-%d')`'_ yearly comparison {.tabset .tabset-fade .tabset-pills}

### Total numbers

```{r}
show.mortality.comp(yearly.comp) +
  labs(title = "Absolute comparison between years (March-15 to {format(Sys.Date(), '%B-%d')})" %>% glue::glue())
```

### Relative numbers _(in percentage)_

```{r}
show.mortality.comp(yearly.comp.rel, is.percentage = TRUE) +
  labs(title = "Relative comparison between years (March-15 to {format(Sys.Date(), '%B-%d')})" %>% glue::glue())
```

## _'March-15 -> `r format(Sys.Date(), '%B-%d')`'_ yearly comparisons with 2009-2019 stats {.tabset .tabset-fade .tabset-pills}

### Total numbers

```{r}
yearly.comp.2 <- sapply(seq(nrow(yearly.stats.2)), function(ix) { yearly.stats.2$sum - yearly.stats.2$sum[ix]}) %>%
  data.frame()
yearly.comp.2[upper.tri(yearly.comp.2)] <- NA
colnames(yearly.comp.2) <- yearly.stats.2$variable
yearly.comp.2 <- bind_cols(year = yearly.stats.2$variable, yearly.comp.2)


show.mortality.comp(yearly.comp.2) +
  labs(title = 'Absolute comparison between pseudo-years', caption = "Max/min/mean/median are calculated day by day in period 2009-2019 (for example, Max is the worse case scenario from that period)\nData retrieved on {format(Sys.Date(), '%B %d, %Y')}" %>% glue::glue()) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 0))
```

### Relative numbers _(in percentage)_

```{r}
show.mortality.comp(yearly.comp.rel.2, is.percentage = TRUE) +
  labs(title = 'Relative comparison between pseudo-years', caption = "Max/min/mean/median are calculated day by day in period 2009-2019 (for example, Max is the worse case scenario from that period)\nData retrieved on {format(Sys.Date(), '%B %d, %Y')}" %>% glue::glue()) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 0))
```

## External causes of death  {.tabset .tabset-fade .tabset-pills}

Accidents, suicides and others. It excludes natural causes or diseases.

### Period of COVID-19

```{r}
external.causes.plot.covid
```

### Full year (`r format(Sys.Date(), '%Y')` is still incomplete)

```{r}
external.causes.plot.full.year
```




















## Mortality by age groups  {.tabset .tabset-fade .tabset-pills}

### Period of COVID-19

```{r}
age.group.plot.covid
```

### Full year (`r format(Sys.Date(), '%Y')` is still incomplete)

```{r}
age.group.plot.full.year
```




```{r, eval=FALSE, include=FALSE}
age.group %>% 
  ggplot() +
  #geom_line(aes(x = day, group = year, color = year)) +
  geom_smooth(aes(x = day, y = `Eventual suicídio`, color = year, group = year), span = 0.1, method = 'loess', formula = y ~ x, na.rm = TRUE, se = FALSE) +
  scale_x_discrete(breaks = function(ix) {
      ix[c(FALSE, TRUE, TRUE, TRUE, TRUE)] <- FALSE
      return(ix)
    }) +
  theme_minimal() +
  theme(legend.key = element_rect(fill = "white", colour = "white"), legend.position = 'top', axis.text.x = element_text(angle = 45, hjust = 1, size = 5))
```


```{r, child='_header.md'}
```